### 2017-04-18 submission/supplemental.Rmd
discussion changes

---
title: ""
bibliography:
csl: microbiome.csl
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header: header.tex
fontsize: 12pt
geometry: margin=1.0in
---

# Differences in the Stool Microbiome Before and After Colorectal Cancer Treatment

\vspace{10mm}

\begin{center}
Marc A Sze, Nielson T Baxter, Mack T Ruffin IV, Mary AM Rogers, and Patrick D Schloss

\vspace{10mm}

Supplemental
\end{center}

\newpage

```{r knitr_settings, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

source('../code/functions.R')

loadLibs(c("tidyr", "dplyr", "scales", "knitr", "rmarkdown"))

```

**Table S1: P-values of Differences between Initial and Follow Up Samples for Richness, Shannon Diversity, and Evenness.**

```{r supplemental_table1, echo=FALSE, message=FALSE, warning=FALSE}

alpha_table <- read.csv("../data/process/tables/alpha_table_summary.csv", header = T, stringsAsFactors = F)
supplemental_table1 <- separate(alpha_table, X, c("comparison", "alpha_metric"), sep="_") %>% 
  mutate(alpha_metric = gsub("sobs", "richness", alpha_metric), 
         comparison = c(rep("Adenoma", 3), rep("Advanced Adenoma", 3), rep("Carcinoma", 3)))

colnames(supplemental_table1) <- c("Comparison", "Alpha Metric", "P-value", "BH Adjusted P-value")

kable(supplemental_table1, format = "markdown", align = 'c', padding = 2)
```

\newpage

\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}

**Table S2: Summary of Common OTUs for All Diagnosis Groups**

\footnotesize

```{r supplemental_table2, echo=FALSE, message=FALSE, warning=FALSE}

common_vars_summary <- read.csv('../data/process/tables/pvalue_adn_srn_crc_common_imp_vars.csv', header = T, row.names = 1, stringsAsFactors = F)

adn_varsMDA <- read.csv('../data/process/tables/adn_reduced_model_top_vars_MDA_Summary.csv', header = T, stringsAsFactors = F)

srn_varsMDA <- read.csv('../data/process/tables/srn_reduced_model_top_vars_MDA_Summary.csv', header = T, stringsAsFactors = F)

crc_varsMDA <- read.csv('../data/process/tables/reduced_crc_model_top_vars_MDA_Summary.csv', header = T, stringsAsFactors = F)

# Create selection vector
common_vars <- common_vars_summary$otu
common_num <- as.numeric(gsub("Otu", "", common_vars_summary$otu))

# Create MDA vectors
adn_MDA <- filter(adn_varsMDA, otu %in% common_vars) %>% slice(match(common_vars, otu)) %>% select(median_MDA)
srn_MDA <- filter(srn_varsMDA, otu %in% common_vars) %>% slice(match(common_vars, otu)) %>% select(median_MDA)
crc_MDA <- filter(crc_varsMDA, otu %in% common_vars) %>% slice(match(common_vars, otu)) %>% select(median_MDA)

st2 <- select(common_vars_summary, otu) %>%
  mutate(Tax = gsub("_unclassified", "", common_vars_summary$lowest_ID)) %>%
  mutate(otu = paste("OTU", common_num, sep = "")) %>%
  mutate(adn_MDA = round(adn_MDA$median_MDA, digits = 2), srn_MDA = round(srn_MDA$median_MDA, digits = 2), crc_MDA = round(crc_MDA$median_MDA, digits = 2)) %>%
  arrange(desc(adn_MDA), desc(srn_MDA), desc(crc_MDA))

colnames(st2) <- c("OTU", "Lowest Taxonomic ID", "Adenoma Median MDA", "Advanced Adenoma Median MDA", "Carcinoma Median MDA")

kable(st2, format = "markdown", align = 'c', padding = 0)
```
\normalsize
\newpage

**Table S3: Summary of Comparisons Based on Additive Treatment Received (Chemotherapy or Chemotherapy and Radiation)**
\footnotesize
```{r supplemental_table3, echo=FALSE, message=FALSE, warning=FALSE}

chemo_rad_summary <- read.csv('../data/process/tables/crc_probs_chemo_rad_pvalue_summary.csv',
                              header = T, stringsAsFactors = F) %>% filter(X != "chemo_v_rads_red") %>% 
  separate(X, c("comparison", "variable"), sep = "_") %>% mutate(comparison = c("Richness", "Shannon", "Evenness", "Thetayc", "Positive Probability")) %>% select(-variable)

st3 <- cbind(chemo_rad_summary$comparison, format(chemo_rad_summary$chemo_rad_mean, digits = 1, nsmall = 1), format(chemo_rad_summary$removal_only_mean, digits = 1, nsmall = 1), format(chemo_rad_summary$cr_sd, digits = 1, nsmall = 1), format(chemo_rad_summary$ro_sd, digits = 1, nsmall = 1), format(chemo_rad_summary$pvalue, scientific = TRUE, digits = 2), format(chemo_rad_summary$bh, scientific = TRUE, digits = 2))

colnames(st3) <- c("Variable", "Additive Treatment Mean", "Removal Mean", "Additive Treatment SD", "Removal SD", "P-value", "BH Corrected")

kable(st3, format = "markdown", align = 'c', padding = 2)

```
\normalsize

\newpage

**Table S4: Summary of Adenoma based on Surgery Received**

```{r supplemental_table4, echo=FALSE, message=FALSE, warning=FALSE}

adn_surg_summary <- read.csv('../data/process/tables/adn_combined_probs_surgery_pvalue_summary.csv',
                              header = T, stringsAsFactors = F) %>%
  filter(X != "prop_surg_adn_srn") %>%
  mutate(variable = c("Richness", "Shannon", "Evenness", "Thetayc", "Adenoma Only\nPositive Probability", "Advanced Adenoma Only\nPositive Probablity"))

st4 <- cbind(adn_surg_summary$variable, format(adn_surg_summary$surg_mean, digits = 1, nsmall = 1), format(adn_surg_summary$no_surg_mean, digits = 1, nsmall = 1), format(adn_surg_summary$surg_sd, digits = 1, nsmall = 1), format(adn_surg_summary$no_surg_sd, digits = 1, nsmall = 1), format(adn_surg_summary$pvalue, scientific = TRUE, digits = 2), format(adn_surg_summary$bh, scientific = TRUE, digits = 2))

colnames(st4) <- c("Variable", "Surgery Mean", "No Surgery Mean", "Surgery SD", "No Surgery SD", "P-value", "BH Corrected")

kable(st4, format = "markdown", align = 'c', padding = 2)

```