### 2017-04-24 submission/manuscript.tex
re-rendering as requested

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}

\usepackage{pdfpages}
\usepackage{comment}
\usepackage{lscape}

\begin{document}

\section{The effect of treatment on the microbiota of patients diagnosed
with colonic
lesions}\label{the-effect-of-treatment-on-the-microbiota-of-patients-diagnosed-with-colonic-lesions}

\begin{center}
\vspace{25mm}

Marc A Sze${^1}$, Nielson T Baxter${^2}$, Mack T Ruffin IV${^3}$, Mary AM Rogers${^2}$, and Patrick D Schloss${^1}$${^\dagger}$

\vspace{20mm}

$\dagger$ To whom correspondence should be addressed: pschloss@umich.edu

$1$ Department of Microbiology and Immunology, University of Michigan, Ann Arbor, MI

$2$ Department of Internal Medicine, University of Michigan, Ann Arbor, MI

$3$ Department of Family Medicine and Community Medicine, Penn State Hershey Medical Center, Hershey, PA

\end{center}

\newpage

\linenumbers

\subsection{Abstract}\label{abstract}

\textbf{Background.} Colorectal cancer (CRC) is a worldwide health
problem. Despite growing evidence that members of the gut microbiota can
drive tumorigenesis, little is known about what happens to the
microbiota after treatment for an adenoma or carcinoma. This study
tested the hypothesis that treatment for adenoma or carcinoma alters the
abundance of bacterial populations associated with disease to those
associated with a normal colon. We tested this hypothesis by sequencing
the 16S rRNA genes in the feces of 67 individuals before and after
treatment for adenoma (N=22), advanced adenoma (N=19), and carcinoma
(N=26).

\textbf{Results.} There were large changes to the bacterial communities
associated with treatment across the three groups. The communities from
patients with carcinomas changed significantly more than those with
adenoma following treatment (P-value \textless{} 0.001); however, there
was not a significant difference between those with advanced adenoma and
those with adenoma or carcinoma (P-value \textgreater{} 0.05). Although
treatment brought about large intrapersonal changes, the change in the
abundance of individual OTUs to treatment was not consistent within
diagnosis groups (P-value \textgreater{} 0.05). Because the distribution
of OTUs across patients and diagnosis groups was patchy, we used the
Random Forest machine learning algorithm to identify groups of OTUs that
allowed us to successfully distinguish between pre and post-treatment
samples for each of the diagnosis groups. Although the three models
successfully differentiated between the pre and post-treatment samples,
there was little overlap between the OTUs that were indicative of
treatment. Next, we used a larger cohort that contained individuals with
normal colons and those with adenomas, advanced adenomas, and carcinomas
to determine whether individuals who underwent a treatment were more
likely to have OTUs associated with normal colons. We again built Random
Forest models and measured the change in the positive probability of
having one of the three diagnoses. Only those patients that initially
had carcinomas experienced a significant decrease in positive
probability of having a lesion (P-value \textless{} 0.05). Finally, we
tested whether the type of treatment impacted the microbiota of those
diagnosed with carcinomas and were unable to detect any significant
differences in characteristics of these communities between individuals
treated with surgery alone and those treated with chemotherapy or
chemotherapy and radiation (P-value \textgreater{} 0.05).

\textbf{Conclusions.} Further exploration of the relationship between
diagnosis, treatment, and the impact on the microbiota will yield
improvements in disease management.

\subsubsection{Keywords}\label{keywords}

microbiota; colorectal cancer; polyps; treatment; risk factor.

\newpage

\linenumbers

\subsection{Background}\label{background}

Colorectal cancer (CRC) is the third most common cause of cancer deaths
in the United States {[}1,2{]}. Disease mortality has significantly
decreased, thanks to improvements in screening {[}1{]}. Despite this
improvement, there are still approximately 50,000 CRC-related deaths per
year {[}2{]}. Current estimates indicate that 20-30\% of those who
undergo treatment will experience recurrence and 30-50\% of those
patients will die {[}3,4{]}. Identification of methods to assess
patients' risk of recurrence is of great importance to reduce mortality
and healthcare costs.

There is growing evidence that the gut microbiota is involved in the
progression of CRC. Mouse-based studies have identified populations of
\emph{Bacteroides fragilis}, \emph{Escherichia coli}, and
\emph{Fusobacterium nucleatum} that alter disease progression
{[}5--9{]}. Furthermore, studies that alter the structure of the
microbiota through the use of antibiotics or inoculation of germ free
mice with human feces has shown that varying community compositions can
result in varied tumor burden {[}10--12{]}. Collectively, these studies
support the hypothesis that the microbiota can alter the amount of
inflammation in the colon and with it the rate of tumorigenesis
{[}13{]}.

Building upon this evidence, several human-based studies have identified
unique signatures of colonic lesions {[}14--19{]}. One line of research
has identified community-level differences between those bacteria that
are found on and adjacent to colonic lesions and have supported a role
for \emph{Bacteroides fragilis}, \emph{Escherichia coli}, and
\emph{Fusobacterium nucleatum} in tumorigenesis {[}20--22{]}. Others
have proposed feces-based biomarkers that could be used to diagnose the
presence of colonic adenomas and carcinomas {[}23--25{]}. These studies
have associated \emph{Fusobacterium nucleatum} and other oral pathogens
with colonic lesions (adenoma, advanced adenoma, and carcinoma). They
have also noted that the loss of bacteria generally thought to produce
short chain fatty acids, which can suppress inflammation, is associated
with colonic lesions. Gut bacteria have a role in tumorigenesis and
there is promise that these populations may be useful biomarkers for
aiding in the early detection of disease.

Despite advances in understanding the role between the gut microbiota
and colonic tumorigenesis, we still do not understand how treatments
including resection, chemotherapy, and/or radiation affect the
composition of the gut microbiota. If the community drives tumorigenesis
then one would hypothesize that treatment to remove a lesion would
affect the microbiota and risk of recurrence. To test this hypothesis,
we addressed two related questions: Does treatment affect the colonic
microbiota in a predictable manner? If so, does the treatment alter the
community to more closely resemble that of individuals with normal
colons?

We answered these questions by sequencing the V4 region of 16S rRNA
genes amplified from fecal samples of individuals with adenoma, advanced
adenoma, and carcinomas pre and post-treatment. We used classical
community analysis to compare the alpha and beta-diversity of
communities pre and post treatment. Next, we generated Random Forest
models to identify bacterial populations that were indicative of
treatment for each diagnosis group. Finally, we measured the predictive
probability of diagnosis models to determine whether treatment would
alter the probability of being diagnosed as having a normal colon or the
colon of the original diagnosis group. We found that treatment alters
the composition of the gut microbiota; however, we only observed a shift
from diseased to normal colon for those initially diagnosed with
carcinomas. Understanding how the community responds to these treatments
could be a valuable tool for identifying biomarkers to quantify the risk
of recurrence.

\newpage

\subsection{Results}\label{results}

\textbf{\emph{Treatment alters the bacterial community structure of
patients diagnosed with colonic lesions.}} Within our 67-person cohort
we tested whether the microbiota of patients with adenoma (N=22),
advanced adenoma (N=19), or carcinoma (N=26) had any broad differences
between pre and post-treatment samples {[}Table 1{]}. The structure of
the microbial communities of the pre and post-treatment samples differed
{[}Figure 1A{]}. We found that the communities obtained pre and
post-treament among the patients with carcinomas changed significantly
more than those patients with adenoma (P-value = 5.4e-05). There were no
significant differences in the amount of change observed between the
patients with adenoma and advanced adenoma or between the patients with
advanced adenoma and carcinoma (P-value \textgreater{} 0.05). Next, we
tested whether there was a consistent direction in the change in the
community structure between the pre and post-treatment samples for each
of the diagnosis groups {[}Figure 1B-D{]}. We only observed a consistent
shift in community structure for the patients with carcinoma when using
a PERMANOVA test (adenoma P-value=0.999, advanced adenoma P-value=0.945,
and carcinoma P-value=0.005). Finally, we measured the number of
observed OTUs, Shannon evenness, and Shannon diversity in the pre and
post-treatment samples and did not observe a significant change for any
of the diagnosis groups (P-value \textgreater{} 0.05) {[}Table S1{]}.

\textbf{\emph{The effects of treatment are not consistent across
treatment groups.}} We used two approaches to identify those bacterial
populations that change between the two samples for each diagnosis
group. First, we sought to identify individual OTUs that could account
for the change in overall community structure. However, using a paired
Wilcoxon test we were unable to identify any OTUs that were
significantly different in the pre and post-treatment groups {[}Figure
S1{]}. It is likely that high inter-individual variation and the patchy
distribution of OTUs across individuals limited the statistical power of
the test. To overcome these problems we developed Random Forest models
to identify collections of OTUs that would allow us to differentiate
between pre and post-treatment samples from each of the diagnosis
groups. Because of the relatively small number of subjects in each
group, we restricted our models to only incorporate 10 OTUs to limit the
likelihood that the models would overfit the data. Despite this
restriction, the models performed well (adenoma AUC=0.69 - 0.92,
advanced adenoma AUC=0.80 - 1.00, carcinoma AUC=0.82 - 0.98).
Interestingly, the 10 OTUs that were used for each model had little
overlap with each other {[}Figure 2{]}. These results support the
earlier community-wide analysis where we observed that the treatment had
an impact on the overall community structure; however, the effect of
treatment was not consistent across patients and diagnosis groups.

\textbf{\emph{Post-treatment samples from patients with carcinoma change
towards a microbiota associated with normal}} Next, we determined
whether treatment changed the microbiota in a way that the
post-treatment communities resembled that of patients with normal
colons. To test this, we used an expanded cohort of 423 individuals that
were diagnosed under the same protocol as having normal colons or colons
with adenoma, advanced adenoma, or carcinoma {[}Table 2{]}. We then
constructed Random Forest models to classify the pre and post-treatment
samples as having their original diagnosis or having a normal colon. The
models performed well (adenoma AUC=0.62 - 0.72, advanced adenoma
AUC=0.68 - 0.77, carcinoma AUC=0.84 - 0.90; Figure S2). The OTUs that
were incorporated into the adenoma and advanced adenoma models largely
overlapped and those OTUs that were used to classify the carcinoma
samples were largely distinct from those of the other two models
{[}Figure 3A{]}. Among the OTUs that were shared across the three models
were those populations commonly considered as commensals (e.g.
\emph{Faecalibacterium}, \emph{Lachnospiraceae}, \emph{Bacteroides},
\emph{Dorea}, \emph{Anaerostipes}, and \emph{Roseburia}) {[}Figures
3B{]}. Although many of these OTUs were also included in the model
differentiating between patients with normal colons and those with
carcinoma, this model also included OTUs affiliated with populations
that have previously been associated with carcinoma
(\emph{Fusobacterium}, \emph{Porphyromonas}, \emph{Parvimonas})
{[}Figure S3{]}. Finally, we applied these three models to the pre and
post-treatment samples for each diagnosis group and quantified the
change in the positive probability of the model. A decrease in the
positive probability would indicate that the colon more closely
resembled that of a patient with a normal colon. There was no
significant change in the positive probability for the adenoma or
advanced adenoma groups {[}Figure 4{]}. The positive probability for the
pre and post-treatment samples from patients diagnosed with carcinoma
significantly decreased with treatment. In fact, 6 of the 26 patients
(23.08\%) that were initially diagnosed with a carcinoma had a higher
positive probability after treatment. One of those was re-diagnosed with
carcinoma on the follow up visit. These results indicate that although
there are changes in the microbiota associated with treatment, those
experienced by the patients in the carcinoma group were the only ones
where the change was directed towards what would be found in a normal
colon.

\textbf{\emph{Difficult to identify effects of specific treatments on
the change in the microbiota.}} The type of treatment that the patients
received varied across diagnosis groups. Those with adenomas and
advanced adenomas received surgical resection (adenoma, N=4; advanced
adenoma, N=4) or polyp removal during colonoscopy (adenoma, N=18;
advanced adenoma, N=15) and those with carcinomas received surgical
resection (N=12), surgical resection with chemotherapy (N=9), and
surgical resection with chemotherapy and radiation (N=5). We focused on
the patients with carcinoma and pooled those patients that received
chemotherapy with those that received chemotherapy and radiation to
improve our statistical power. We did not observe a significant
difference in the effect of these treatments on the number of observed
OTUs, Shannon diversity, or Shannon evenness (P-value \textgreater{}
0.05). Furthermore, there was not a significant difference in the effect
of the treatments on the amount of change in the community structure (P=
1). Finally, the change in the positive probability was not
significantly different between the two treatment groups (P=1). Due to
the relatively small number of samples in each treatment group, it was
difficult to make a definitive statement regarding the specific type of
treatment on the amount of change in the structure of the microbiota.

\newpage

\subsection{Discussion}\label{discussion}

This study builds upon previous work that demonstrated a role for the
microbiota in tumorigenesis and that the microbiota contains biomarkers
to diagnosis colonic lesions {[}5--9,23--27{]}. Our study focused on
comparing the microbiota of patients diagnosed with adenoma, advanced
adenoma, and carcinoma before and after treatment. For all three groups
of patients we observed changes in their microbiota. The microbiota of
patients diagnosed with carcinoma changed significantly more than the
other groups. This change resulted in communities that more closely
resembled those of patients with a normal colon. For the patients
diagnosed with carcinomas there was a larger change between their
initial and follow up samples than was observed for patients diagnosed
with either adenoma or advanced adenoma. The changes in the community
structure of those diagnosed with carcinoma were similar and resulted in
communities that more closely resembled those found in patients with
normal colons.

Understanding the effect of treatment on the microbiota of those
initially diagnosed with carcinomas may have important implications into
understanding disease recurrence. The microbiota changes associated with
treatment among patients initially diagnosed with carcinomas resulted in
community structures that resembled patients with normal colons. This
would suggest that treatment for carcinoma is not only successful for
removing the carcinoma but also at removing bacteria associated with
them. Interestingly, 6 of the 26 patients diagnosed with carcinomas had
an elevated probability of having carcinomas after treatment. This leads
us to hypothesize that these individuals may be at a high risk of
recurrence. Interestingly, the 23.08\% prevalence of increased carcinoma
probability from our study is within the expected rate of recurrence
(20-30\% {[}3,4{]}). Unfortunately, because of the design of our study,
it was not possible to determine whether individuals who had an
increased positive probability for carcinoma experienced recurrence.
Regardless, it is intriguing that it may be possible to use
microbiome-based biomarkers to not only predict the presence of lesions,
but to also assess the risk of developing them.

It is interesting that those patients diagnosed with adenoma and
advanced adenoma did not experience a shift towards a community
structure that resembled those with normal colons. This may be due to
the fundamental differences between the features of adenomas and
advanced adenomas and carcinoma. Specifically, carcinomas may create an
inflammatory milieu that would impact the structure of the community and
removal of that stimulus would alter said structure. In addition, it is
possible that the difference between the microbiota of patients with
adenoma and advanced adenoma and those with normal colons is subtle.
This is supported by the reduced ability of our models to correctly
classify patients with adenomas and advanced adenomas relative to those
diagnosed with carcinomas {[}Figure S2{]}. Given the patchy distribution
of microbiota across patients in the different diagnosis groups, it is
possible that we lacked the statistical power to adequately characterize
the change in the communities following treatment. A final hypothesis is
that the specific type of treatment altered the structure of the
microbiome. The treatment to remove adenomas and advanced adenomas was
either polyp removal or surgical resection whereas it was surgical
resection alone or in combination with chemotherapy or with chemotherapy
and radiation for individuals with carcinoma. Because chemotherapy and
radiation target rapidly growing cells, these treatments would be more
likely to cause a turn over of the colonic epithelium driving a more
significant change in the structure of the microbiota. Although, we were
able to test for an effect across these specific types of treatment, the
number of patients in each treatment group was relatively small.

It was not clear why 6 of the 26 patients with carcinoma had a higher
probability of having carcinomas after treatment. We hypothesized that
these individuals may have had more severe tumors; however, the tumor
severity of these 6 individuals (3 with Stage II and 3 with Stage III)
was similar to the distribution observed among the other 20 patients. We
also hypothesized that we may have sampled these patients later than the
rest and their communities may have reverted to a carcinoma-associated
state; however, there was not a statistically significant difference in
the length of time between sample collection among those whose
probabilities increased or decreased (Wilcoxon Test; P-value = 0.56).
Finally, it is possible that these patients may not have responded to
treatment as well as the other 20 patients diagnosed with carcinoma and
so the microbiota may not have been impacted the same way. Again,
further studies looking at the role of the microbiota in recurrence are
needed to understand the dynamics following treatment.

This study builds upon existing work that has established a role for the
microbiota in tumorigenesis and that demonstrated the utility of
microbiome-based biomarkers to predict the presence of colonic lesions.
The most exciting future direction from the current study is the
possibility that markers within the microbiota could be used to evaluate
the effect of treatment and predict recurrence for those initially
diagnosed with carcinoma. If such an approach is effective, it might be
possible to even target the microbiota as part of combined therapies
including currently standard approaches. Our data provides additional
evidence on the importance of the microbiota in tumorigenesis by
addressing the recovery of the microbiota after treatment and opens
interesting avenues of research into how these changes may affect
recurrence.

\newpage

\subsection{Methods}\label{methods}

\textbf{\emph{Study Design and Patient Sampling.}} Sampling and design
have been previously reported in Baxter, et al {[}23{]}. Briefly, study
exclusion involved those who had already undergone surgery, radiation,
or chemotherapy, had colorectal cancer before a baseline fecal sample
could be obtained, had IBD, a known hereditary non-polyposis colorectal
cancer, or familial adenomatous polyposis. Samples used to build the
models for prediction were collected either prior to a colonoscopy or
between one and two weeks after initial colonoscopy. The bacterial
community has been shown to normalize back to a pre-colonoscopy
community within this time period {[}28{]}. Our study cohort consisted
of 67 individuals with an initial sample as described and a follow up
sample obtained between 188 - 546 days after treatment of lesion
{[}Table 1{]}. Patients were diagnosed by colonoscopic examination and
histopathological review of biopsies. Patients were classified as having
advanced adenoma if they had an adenoma greater than 1 cm, more than
three adenomas of any size, or an adenoma with villous histology. This
study was approved by the University of Michigan Institutional Review
Board. All study participants provided informed consent and the study
itself conformed to the guidelines set out by the Helsinki Declaration.

\textbf{\emph{16S rRNA Gene Sequencing.}} Sequencing was completed as
described by Kozich, et al. {[}29{]}. DNA extraction used the 96-well
Soil DNA isolation kit (MO BIO Laboratories) and an epMotion 5075
automated pipetting system (Eppendorf). The V4 variable region was
amplified and the resulting product was split between three sequencing
runs with normal, adenoma, and carcinoma evenly represented on each run.
Each group was randomly assigned to avoid biases based on sample
collection location. The pre and post-treatment samples were sequenced
on the same run.

\textbf{\emph{Sequence Processing.}} The mothur software package
(v1.37.5) was used to process the 16S rRNA gene sequences and has been
previously described {[}29{]}. The general workflow using mothur
included merging paired-end reads into contigs, filtering for low
quality contigs, aligning to the SILVA database {[}30{]}, screening for
chimeras using UCHIME {[}31{]}, classifying with a naive Bayesian
classifier using the Ribosomal Database Project (RDP){[}32{]}, and
clustered into Operational Taxonomic Units (OTUs) using a 97\%
similarity cutoff with an average neighbor clustering algorithm
{[}33{]}. The number of sequences for each sample was rarefied to 10523
to minimize the impacts of uneven sampling.

\textbf{\emph{Model Building.}} The Random Forest {[}34{]} algorithm was
used to create the model used to create the three models used. The
adenoma model classified normal versus adenoma, advanced adenoma was
normal versus advanced adenoma, and carcinoma was normal versus
carcinoma. The toal number of individuals in this data set was 423
individuals {[}Table 1{]}. There were a total of 239 individuals in the
adenoma model, 262 individuals in the advanced adenoma model, and 266
indivdiuals in the carcinoma model {[}Table 1{]}. Each model was then
applied to our 67-person cohort testing prediction of adenoma intial
(adenoma n = 22) versus adenoma follow up (adenoma n = 0), advanced
adenoma initial (advanced adenoma n = 19) versus advanced adenoma follow
up (advanced adenoma n = 0), carcinoma initial (carinoma n = 26) versus
carcinoma follow up (carcinoma n = 1).

The model included only OTU data obtained from 16S rRNA sequencing.
Non-binary data was checked for near zero variance and OTUs that had
near zero variance were removed. This pre-processing was performed with
the R package caret (v6.0.73). Optimization of the mtry hyper-parameter
involved making 100 different 80/20 (train/test) splits of the data
where normal and adenoma, normal and advanced adenoma, or normal and
carcinoma were represented in the same proportion within both the whole
data set and the 80/20 split. For each of the different splits, 20
repeated 10-fold cross validation was performed on the 80\% component to
optimize the mtry hyper-parameter by maximizing the AUC (Area Under the
Curve of the Receiver Operator Characteristic). The resulting model was
then tested on the hold out data obtained from the 20\% component. All
three models had an optimized mtry of 2.

Assessment of the most important OTUs to the model involved counting the
number of times an OTU was present in the top 10\% of mean decrease in
accuracy (MDA) for each of the 100 different splits run. This was then
followed with filtering of this list to variables that were only present
in more than 50\% of these 100 runs. The final collated list of
variables was then run through the mtry optimization again. Once the
ideal mtry was found the entire sample set specific to normal versus
adenoma, normal versus advanced adenoma, or normal versus carcinoma was
used to create the final Random Forest model on which classifications on
the 67-person cohort was completed. For all three models the final
optimized mtry was 2. The hyper-parameter, mtry, defines the number of
variables to investigate at each split before a new division of the data
was created with the Random Forest model.

\textbf{\emph{Statistical Analysis.}} The R software package (v3.3.2)
was used for all statistical analysis. Comparisons between bacterial
community structure utilized PERMANOVA {[}35{]} in the vegan package
(v2.4.1). Comparisons between probabilities as well as overall OTU
differences between pre and post-treatment samples utilized a paired
Wilcoxon ranked sum test. Where multiple comparison testing was
appropriate, a Benjamini-Hochberg (BH) correction was applied {[}36{]}
and a corrected P-value of less than 0.05 was considered significant.
Unless otherwise stated the P-values reported are those that were BH
corrected. Model rank importance was determined by obtaining the median
MDA from the 100, 20 repeated 10-fold cross validation and then ranking
from largest to smallest MDA.

\textbf{\emph{Reproducible Methods.}} A detailed and reproducible
description of how the data were processed and analyzed can be found at
\url{https://github.com/SchlossLab/Sze_followUps_2017}. Raw sequences
have been deposited into the NCBI Sequence Read Archive (SRP062005 and
SRP096978) and the necessary metadata can be found at
\url{https://www.ncbi.nlm.nih.gov/Traces/study/} and searching the
respective SRA study accession.

\newpage

\textbf{Figure 1: General Differences between Adenoma, Advanced Adenoma,
and Carcinoma Groups After Treatment.} A) A significant difference was
found between the adenoma and carcinoma group for thetayc (P-value =
NULL). Solid black points represent the median value for each diagnosis
group. B) NMDS of the pre and post-treatment samples for the adenoma
group. C) NMDS of the pre and post-treatment samples for the advanced
adenoma group. D) NMDS of the pre and post-treatment samples for the
carcinoma group.

\textbf{Figure 2: The 10 OTUs used to classify treatment for Adenoma,
Advanced Adenoma, and Carcinoma.} A) Adenoma OTUs. B) Advanced Adenoma
OTUs. C) Carcinoma OTUs.

\textbf{Figure 3: Treatment Response Based on Models Built for Adenoma,
Advanced Adenoma, or Carcinoma.} A) Positive probability change from
initial to follow up sample in those with adenoma. B) Positive
probability change from initial to follow up sample in those with
advanced adenoma. C) Positive probability change from initial to follow
up sample in those with carcinoma..

\textbf{Figure 4: Common OTUs to All Models.} A) Venn diagram showing
the OTU overlap between each model. B) For each common OTU the lowest
taxonomic identification and importance rank for each model run is
shown.

\newpage

\textbf{Table 1: Demographic Data of pre and Post Treatment Cohort}

\textbf{Table 2: Demographic Data of Training Cohort}

\newpage

\textbf{Figure S1: Distribution of P-values from Paired Wilcoxson
Analysis of All OTUs Before and After Treatment}

\textbf{Figure S2: ROC Curves of the Adenoma, Advanced Adenoma, and
Carcinoma Models.} A) Adenoma ROC curve: The light green shaded areas
represent the range of values of a 100 different 80/20 splits of the
test set data and the dark green line represents the model using 100\%
of the data set and what was used for subsequent classification. B)
Advanced Adenoma ROC curve: The light yellow shaded areas represent the
range of values of a 100 different 80/20 splits of the test set data and
the dark yellow line represents the model using 100\% of the data set
and what was used for subsequent classification. C) Carcinoma ROC curve:
The light red shaded areas represent the range of values of a 100
different 80/20 splits of the test set data and the dark red line
represents the model using 100\% of the data set and what was used for
subsequent classification.

\textbf{Figure S3: Summary of Important OTUs for the Adenoma, Advanced
Adenoma, and Carcinoma Models.} A) MDA of the most important variables
in the adenoma model. The dark green point represents the mean and the
lighter green points are the value of each of the 100 different runs. B)
Summary of Important Variables in the advanced adenoma model. MDA of the
most important variables in the SRN model. The dark yellow point
represents the mean and the lighter yellow points are the value of each
of the 100 different runs. C) MDA of the most important variables in the
carcinoma model. The dark red point represents the mean and the lighter
red points are the value of each of the 100 different runs.

\newpage

\subsection{Declarations}\label{declarations}

\subsubsection{Ethics approval and consent to
participate}\label{ethics-approval-and-consent-to-participate}

The University of Michigan Institutional Review Board approved this
study, and all subjects provided informed consent. This study conformed
to the guidelines of the Helsinki Declaration.

\subsubsection{Consent for publication}\label{consent-for-publication}

Not applicable.

\subsubsection{Availability of data and
material}\label{availability-of-data-and-material}

A detailed and reproducible description of how the data were processed
and analyzed can be found at
\url{https://github.com/SchlossLab/Sze_followUps_2017}. Raw sequences
have been deposited into the NCBI Sequence Read Archive (SRP062005 and
SRP096978) and the necessary metadata can be found at
\url{https://www.ncbi.nlm.nih.gov/Traces/study/} and searching the
respective SRA study accession.

\subsubsection{Competing Interests}\label{competing-interests}

All authors declare that they do not have any relevant competing
interests to report.

\subsubsection{Funding}\label{funding}

This study was supported by funding from the National Institutes of
Health to P. Schloss (R01GM099514, P30DK034933) and to the Early
Detection Research Network (U01CA86400).

\subsubsection{Authors' contributions}\label{authors-contributions}

All authors were involved in the conception and design of the study. MAS
analyzed the data. NTB processed samples and analyzed the data. All
authors interpreted the data. MAS and PDS wrote the manuscript. All
authors reviewed and revised the manuscript. All authors read and
approved the final manuscript.

\subsubsection{Acknowledgements}\label{acknowledgements}

The authors thank the Great Lakes-New England Early Detection Research
Network for providing the fecal samples that were used in this study. We
would also like to thank Amanda Elmore for reviewing and correcting code
error and providing feedback on manuscript drafts. We would also like to
thank Nicholas Lesniak for providing feedback on manuscript drafts.

\newpage

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\hypertarget{ref-jemal_cancer_2010}{}
1. Jemal A, Siegel R, Xu J, Ward E. Cancer statistics, 2010. CA: a
cancer journal for clinicians. 2010;60:277--300.

\hypertarget{ref-haggar_colorectal_2009}{}
2. Haggar FA, Boushey RP. Colorectal cancer epidemiology: Incidence,
mortality, survival, and risk factors. Clinics in Colon and Rectal
Surgery. 2009;22:191--7.

\hypertarget{ref-hellinger_reoperation_2006}{}
3. Hellinger MD, Santiago CA. Reoperation for recurrent colorectal
cancer. Clinics in Colon and Rectal Surgery. 2006;19:228--36.

\hypertarget{ref-ryuk_predictive_2014}{}
4. Ryuk JP, Choi G-S, Park JS, Kim HJ, Park SY, Yoon GS, et al.
Predictive factors and the prognosis of recurrence of colorectal cancer
within 2 years after curative resection. Annals of Surgical Treatment
and Research. 2014;86:143--51.

\hypertarget{ref-goodwin_polyamine_2011}{}
5. Goodwin AC, Destefano Shields CE, Wu S, Huso DL, Wu X, Murray-Stewart
TR, et al. Polyamine catabolism contributes to enterotoxigenic
Bacteroides fragilis-induced colon tumorigenesis. Proceedings of the
National Academy of Sciences of the United States of America.
2011;108:15354--9.

\hypertarget{ref-abed_fap2_2016}{}
6. Abed J, Emgård JEM, Zamir G, Faroja M, Almogy G, Grenov A, et al.
Fap2 Mediates Fusobacterium nucleatum Colorectal Adenocarcinoma
Enrichment by Binding to Tumor-Expressed Gal-GalNAc. Cell Host \&
Microbe. 2016;20:215--25.

\hypertarget{ref-arthur_microbial_2014}{}
7. Arthur JC, Gharaibeh RZ, Mühlbauer M, Perez-Chanona E, Uronis JM,
McCafferty J, et al. Microbial genomic analysis reveals the essential
role of inflammation in bacteria-induced colorectal cancer. Nature
Communications. 2014;5:4724.

\hypertarget{ref-kostic_fusobacterium_2013}{}
8. Kostic AD, Chun E, Robertson L, Glickman JN, Gallini CA, Michaud M,
et al. Fusobacterium nucleatum potentiates intestinal tumorigenesis and
modulates the tumor-immune microenvironment. Cell Host \& Microbe.
2013;14:207--15.

\hypertarget{ref-wu_human_2009}{}
9. Wu S, Rhee K-J, Albesiano E, Rabizadeh S, Wu X, Yen H-R, et al. A
human colonic commensal promotes colon tumorigenesis via activation of T
helper type 17 T cell responses. Nature Medicine. 2009;15:1016--22.

\hypertarget{ref-zackular_manipulation_2016}{}
10. Zackular JP, Baxter NT, Chen GY, Schloss PD. Manipulation of the Gut
Microbiota Reveals Role in Colon Tumorigenesis. mSphere. 2016;1.

\hypertarget{ref-zackular_gut_2013}{}
11. Zackular JP, Baxter NT, Iverson KD, Sadler WD, Petrosino JF, Chen
GY, et al. The gut microbiome modulates colon tumorigenesis. mBio.
2013;4:e00692--00613.

\hypertarget{ref-baxter_structure_2014}{}
12. Baxter NT, Zackular JP, Chen GY, Schloss PD. Structure of the gut
microbiome following colonization with human feces determines colonic
tumor burden. Microbiome. 2014;2:20.

\hypertarget{ref-flynn_metabolic_2016}{}
13. Flynn KJ, Baxter NT, Schloss PD. Metabolic and Community Synergy of
Oral Bacteria in Colorectal Cancer. mSphere. 2016;1.

\hypertarget{ref-wang_structural_2012}{}
14. Wang T, Cai G, Qiu Y, Fei N, Zhang M, Pang X, et al. Structural
segregation of gut microbiota between colorectal cancer patients and
healthy volunteers. The ISME journal. 2012;6:320--9.

\hypertarget{ref-chen_decreased_2013}{}
15. Chen H-M, Yu Y-N, Wang J-L, Lin Y-W, Kong X, Yang C-Q, et al.
Decreased dietary fiber intake and structural alteration of gut
microbiota in patients with advanced colorectal adenoma. The American
Journal of Clinical Nutrition. 2013;97:1044--52.

\hypertarget{ref-chen_human_2012}{}
16. Chen W, Liu F, Ling Z, Tong X, Xiang C. Human intestinal lumen and
mucosa-associated microbiota in patients with colorectal cancer. PloS
One. 2012;7:e39743.

\hypertarget{ref-shen_molecular_2010}{}
17. Shen XJ, Rawls JF, Randall T, Burcal L, Mpande CN, Jenkins N, et al.
Molecular characterization of mucosal adherent bacteria and associations
with colorectal adenomas. Gut Microbes. 2010;1:138--47.

\hypertarget{ref-kostic_genomic_2012}{}
18. Kostic AD, Gevers D, Pedamallu CS, Michaud M, Duke F, Earl AM, et
al. Genomic analysis identifies association of Fusobacterium with
colorectal carcinoma. Genome Research. 2012;22:292--8.

\hypertarget{ref-feng_gut_2015}{}
19. Feng Q, Liang S, Jia H, Stadlmayr A, Tang L, Lan Z, et al. Gut
microbiome development along the colorectal adenoma-carcinoma sequence.
Nature Communications. 2015;6:6528.

\hypertarget{ref-dejea_microbiota_2014}{}
20. Dejea CM, Wick EC, Hechenbleikner EM, White JR, Mark Welch JL,
Rossetti BJ, et al. Microbiota organization is a distinct feature of
proximal colorectal cancers. Proceedings of the National Academy of
Sciences of the United States of America. 2014;111:18321--6.

\hypertarget{ref-mima_fusobacterium_2015}{}
21. Mima K, Sukawa Y, Nishihara R, Qian ZR, Yamauchi M, Inamura K, et
al. Fusobacterium nucleatum and T Cells in Colorectal Carcinoma. JAMA
oncology. 2015;1:653--61.

\hypertarget{ref-arthur_intestinal_2012}{}
22. Arthur JC, Perez-Chanona E, Mühlbauer M, Tomkovich S, Uronis JM, Fan
T-J, et al. Intestinal inflammation targets cancer-inducing activity of
the microbiota. Science (New York, N.Y.). 2012;338:120--3.

\hypertarget{ref-baxter_microbiota-based_2016}{}
23. Baxter NT, Ruffin MT, Rogers MAM, Schloss PD. Microbiota-based model
improves the sensitivity of fecal immunochemical test for detecting
colonic lesions. Genome Medicine. 2016;8:37.

\hypertarget{ref-zeller_potential_2014}{}
24. Zeller G, Tap J, Voigt AY, Sunagawa S, Kultima JR, Costea PI, et al.
Potential of fecal microbiota for early-stage detection of colorectal
cancer. Molecular Systems Biology. 2014;10:766.

\hypertarget{ref-zackular_human_2014}{}
25. Zackular JP, Rogers MAM, Ruffin MT, Schloss PD. The human gut
microbiome as a screening tool for colorectal cancer. Cancer Prevention
Research (Philadelphia, Pa.). 2014;7:1112--21.

\hypertarget{ref-yu_metagenomic_2017}{}
26. Yu J, Feng Q, Wong SH, Zhang D, Liang QY, Qin Y, et al. Metagenomic
analysis of faecal microbiome as a tool towards targeted non-invasive
biomarkers for colorectal cancer. Gut. 2017;66:70--8.

\hypertarget{ref-warren_co-occurrence_2013}{}
27. Warren RL, Freeman DJ, Pleasance S, Watson P, Moore RA, Cochrane K,
et al. Co-occurrence of anaerobic bacteria in colorectal carcinomas.
Microbiome. 2013;1:16.

\hypertarget{ref-obrien_impact_2013}{}
28. O'Brien CL, Allison GE, Grimpen F, Pavli P. Impact of colonoscopy
bowel preparation on intestinal microbiota. PloS One. 2013;8:e62815.

\hypertarget{ref-kozich_development_2013}{}
29. Kozich JJ, Westcott SL, Baxter NT, Highlander SK, Schloss PD.
Development of a dual-index sequencing strategy and curation pipeline
for analyzing amplicon sequence data on the MiSeq Illumina sequencing
platform. Applied and Environmental Microbiology. 2013;79:5112--20.

\hypertarget{ref-pruesse_silva_2007}{}
30. Pruesse E, Quast C, Knittel K, Fuchs BM, Ludwig W, Peplies J, et al.
SILVA: A comprehensive online resource for quality checked and aligned
ribosomal RNA sequence data compatible with ARB. Nucleic Acids Research.
2007;35:7188--96.

\hypertarget{ref-edgar_uchime_2011}{}
31. Edgar RC, Haas BJ, Clemente JC, Quince C, Knight R. UCHIME improves
sensitivity and speed of chimera detection. Bioinformatics (Oxford,
England). 2011;27:2194--200.

\hypertarget{ref-wang_naive_2007}{}
32. Wang Q, Garrity GM, Tiedje JM, Cole JR. Naive Bayesian classifier
for rapid assignment of rRNA sequences into the new bacterial taxonomy.
Applied and Environmental Microbiology. 2007;73:5261--7.

\hypertarget{ref-schloss_assessing_2011}{}
33. Schloss PD, Westcott SL. Assessing and improving methods used in
operational taxonomic unit-based approaches for 16S rRNA gene sequence
analysis. Applied and Environmental Microbiology. 2011;77:3219--26.

\hypertarget{ref-breiman_random_2001}{}
34. Breiman L. Random Forests. Machine Learning {[}Internet{]}. 2001
{[}cited 2013 Feb 7{]};45:5--32. Available from:
\href{http://link.springer.com/article/10.1023/A\%3A1010933404324\%20http://link.springer.com/article/10.1023\%2FA\%3A1010933404324?LI=true}{http://link.springer.com/article/10.1023/A\%3A1010933404324 http://link.springer.com/article/10.1023\%2FA\%3A1010933404324?LI=true}

\hypertarget{ref-anderson_permanova_2013}{}
35. Anderson MJ, Walsh DCI. PERMANOVA, ANOSIM, and the Mantel test in
the face of heterogeneous dispersions: What null hypothesis are you
testing? Ecological Monographs {[}Internet{]}. 2013 {[}cited 2017 Jan
5{]};83:557--74. Available from:
\url{http://doi.wiley.com/10.1890/12-2010.1}

\hypertarget{ref-benjamini_controlling_1995}{}
36. Benjamini Y, Hochberg Y. Controlling the false discovery rate: A
practical and powerful approach to multiple testing. Journal of the
Royal Statistical Society. Series B (Methodological). 1995;57:289--300.

\end{document}