### 2017-04-28 submission/manuscript.tex

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}

\usepackage{pdfpages}
\usepackage{comment}
\usepackage{lscape}

\begin{document}

\section{The effect of treatment on the microbiota of patients diagnosed
with colonic
lesions}\label{the-effect-of-treatment-on-the-microbiota-of-patients-diagnosed-with-colonic-lesions}

\begin{center}
\vspace{25mm}

Marc A Sze${^1}$, Nielson T Baxter${^2}$, Mack T Ruffin IV${^3}$, Mary AM Rogers${^2}$, and Patrick D Schloss${^1}$${^\dagger}$

\vspace{20mm}

$\dagger$ To whom correspondence should be addressed: pschloss@umich.edu

$1$ Department of Microbiology and Immunology, University of Michigan, Ann Arbor, MI

$2$ Department of Internal Medicine, University of Michigan, Ann Arbor, MI

$3$ Department of Family Medicine and Community Medicine, Penn State Hershey Medical Center, Hershey, PA

\end{center}

\newpage

\linenumbers

\subsection{Abstract}\label{abstract}

\textbf{Background.} Colorectal cancer (CRC) is a worldwide health
problem. Despite growing evidence that members of the gut microbiota can
drive tumorigenesis, little is known about what happens to the
microbiota after treatment for an adenoma or carcinoma. This study
tested the hypothesis that treatment for adenoma or carcinoma alters the
abundance of bacterial populations associated with disease to those
associated with a normal colon. We tested this hypothesis by sequencing
the 16S rRNA genes in the feces of 67 individuals before and after
treatment for adenoma (N=22), advanced adenoma (N=19), and carcinoma
(N=26).

\textbf{Results.} There were large changes to the bacterial communities
associated with treatment across the three groups. The communities from
patients with carcinomas changed significantly more than those with
adenoma following treatment (P-value \textless{} 0.001). There was no
significant change in the microbiota between patients with adenoma and
advanced adenoma, or between patients with advanced adenoma and
carcinoma (P-value \textgreater{} 0.05). Although treatment was
associated with intrapersonal changes, the change in the abundance of
individual OTUs to treatment was not consistent within diagnosis groups
(P-value \textgreater{} 0.05). Because the distribution of OTUs across
patients and diagnosis groups was irregular, we used the Random Forest
machine learning algorithm to identify groups of OTUs that allowed us to
successfully distinguish between pre and post-treatment samples for each
of the diagnosis groups. Although the three models successfully
differentiated between the pre and post-treatment samples, there was
little overlap between the OTUs that were indicative of treatment. Next,
we used a larger cohort that contained individuals with normal colons
and those with adenomas, advanced adenomas, and carcinomas to determine
whether individuals who underwent treatment were more likely to have
OTUs associated with normal colons. We again built Random Forest models
and measured the change in the positive probability of having one of the
three diagnoses. Only patients who had carcinomas experienced a
significant decrease in positive probability of having a lesion after
treatment (P-value \textless{} 0.05), indicating that the microbial
milieu of the colon more closely resembled that of a normal colon.
Finally, we tested whether the type of treatment impacted the microbiota
of those diagnosed with carcinomas and were unable to detect any
significant differences in characteristics of these communities between
individuals treated with surgery alone and those treated with
chemotherapy or chemotherapy and radiation (P-value \textgreater{}
0.05).

\textbf{Conclusions.} Further exploration of the relationship between
diagnosis, treatment, and the impact on the microbiota will yield
improvements in disease management.

\subsubsection{Keywords}\label{keywords}

microbiota; colorectal cancer; polyps; treatment; risk factor.

\newpage

\linenumbers

\subsection{Background}\label{background}

Colorectal cancer (CRC) is the third most common cause of cancer deaths
in the United States {[}1,2{]}. Disease mortality has significantly
decreased, predominately due to improvements in screening {[}2{]}.
Despite this improvement, there are still approximately 50,000
CRC-related deaths per year {[}1{]}. Current estimates indicate that
20-30\% of those who undergo treatment will experience recurrence and
35\% of all patients will die within five years {[}3--5{]}.
Identification of methods to assess patients' risk of recurrence is of
great importance to reduce mortality and healthcare costs.

There is growing evidence that the gut microbiota is involved in the
progression of CRC. Mouse-based studies have identified populations of
\emph{Bacteroides fragilis}, \emph{Escherichia coli}, and
\emph{Fusobacterium nucleatum} that alter disease progression
{[}6--10{]}. Furthermore, studies that shift the structure of the
microbiota through the use of antibiotics or inoculation of germ free
mice with human feces have shown that varying community compositions can
result in varied tumor burden {[}11--13{]}. Collectively, these studies
support the hypothesis that the microbiota can alter the amount of
inflammation in the colon and with it the rate of tumorigenesis
{[}14{]}.

Building upon this evidence, several human studies have identified
unique signatures of colonic lesions {[}15--20{]}. One line of research
has identified community-level differences between those bacteria that
are found on and adjacent to colonic lesions and have supported a role
for \emph{Bacteroides fragilis}, \emph{Escherichia coli}, and
\emph{Fusobacterium nucleatum} in tumorigenesis {[}21--23{]}. Others
have proposed feces-based biomarkers that could be used to diagnose the
presence of colonic adenomas and carcinomas {[}24--26{]}. These studies
have associated \emph{Fusobacterium nucleatum} and other oral pathogens
with colonic lesions (adenoma, advanced adenoma, and carcinoma). They
have also noted that the loss of bacteria generally thought to produce
short chain fatty acids, which can suppress inflammation, is associated
with colonic lesions. This suggests that gut bacteria have a role in
tumorigenesis with potential as useful biomarkers for aiding in the
early detection of disease.

Despite advances in understanding the role between the gut microbiota
and colonic tumorigenesis, we still do not understand how treatments
including resection, chemotherapy, and/or radiation affect the
composition of the gut microbiota. If the community drives tumorigenesis
then one would hypothesize that treatment to remove a lesion would
affect the microbiota and risk of recurrence. To test this hypothesis,
we addressed two related questions: Does treatment affect the colonic
microbiota in a predictable manner? If so, does the treatment alter the
community to more closely resemble that of individuals with normal
colons?

We answered these questions by sequencing the V4 region of 16S rRNA
genes amplified from fecal samples of individuals with adenoma, advanced
adenoma, and carcinomas pre and post-treatment. We used classical
community analysis to compare the alpha and beta-diversity of
communities pre and post treatment. Next, we generated Random Forest
models to identify bacterial populations that were indicative of
treatment for each diagnosis group. Finally, we measured the predictive
probabilities to assess whether treatment yielded bacterial communities
similar to those individuals with normal colons. We found that treatment
alters the composition of the gut microbiota and that, for those with
carinomas, the gut microbiota shifted more towards that of a normal
colon after treatment. In the individuals with carcinomas, no difference
was found byt the type of treatment (surgery alone versus surgery with
chemotherapy). Understanding how the community responds to these
treatments could be a valuable tool for identifying biomarkers to
quantify the risk of recurrence and the likelihood of survival.

\newpage

\subsection{Results}\label{results}

\textbf{\emph{Treatment alters the bacterial community structure of
patients diagnosed with colonic lesions.}} Within our 67-person cohort
we tested whether the microbiota of patients with adenoma (N=22),
advanced adenoma (N=19), or carcinoma (N=26) had any broad differences
between pre and post-treatment samples {[}Table 1{]}. The structure of
the microbial communities of the pre and post-treatment samples differed
{[}Figure 1A{]}. We found that the communities obtained pre and
post-treament among the patients with carcinomas changed significantly
more than those patients with adenoma (P-value = 5.4e-05). There were no
significant differences in the amount of change observed between the
patients with adenoma and advanced adenoma or between the patients with
advanced adenoma and carcinoma (P-value \textgreater{} 0.05). Next, we
tested whether there was a consistent direction in the change in the
community structure between the pre and post-treatment samples for each
of the diagnosis groups {[}Figure 1B-D{]}. We only observed a consistent
shift in community structure for the patients with carcinoma when using
a PERMANOVA test (adenoma P-value=0.999, advanced adenoma P-value=0.945,
and carcinoma P-value=0.005). Finally, we measured the number of
observed OTUs, Shannon evenness, and Shannon diversity in the pre and
post-treatment samples and did not observe a significant change for any
of the diagnosis groups (P-value \textgreater{} 0.05) {[}Table S1{]}.

\textbf{\emph{The effects of treatment are not consistent across
treatment groups.}} We used two approaches to identify those bacterial
populations that change between the two samples for each diagnosis
group. First, we sought to identify individual OTUs that could account
for the change in overall community structure. However, using a paired
Wilcoxon test we were unable to identify any OTUs that were
significantly different in the pre and post-treatment groups {[}Figure
S1{]}. It is likely that high inter-individual variation and the
irregular distribution of OTUs across individuals limited the
statistical power of the test. To overcome these problems we developed
Random Forest models to identify collections of OTUs that would allow us
to differentiate between pre and post-treatment samples from each of the
diagnosis groups. Because of the relatively small number of subjects in
each group, we restricted our models to only incorporate 10 OTUs to
limit the likelihood that the models would overfit the data. Despite
this restriction, the models performed well (adenoma AUC=0.69 - 0.92,
advanced adenoma AUC=0.80 - 1.00, carcinoma AUC=0.82 - 0.98).
Interestingly, the 10 OTUs that were used for each model had little
overlap with each other {[}Figure 2{]}. These results support the
earlier community-wide analysis where we observed that the treatment had
an impact on the overall community structure; however, the effect of
treatment was not consistent across patients and diagnosis groups.

\textbf{\emph{Post-treatment samples from patients with carcinoma more
closely resemble those of a normal colon.}} Next, we determined whether
treatment changed the microbiota in a way that the post-treatment
communities resembled that of patients with normal colons. To test this,
we used an expanded cohort of 423 individuals that were diagnosed under
the same protocol as having normal colons or colons with adenoma,
advanced adenoma, or carcinoma {[}Table 2{]}. We then constructed Random
Forest models to classify the study samples, with the 3 diagnosis groups
(adenoma, advanced adenoma, or carcinoma), or having a normal colon. The
models performed well (adenoma AUC=0.62 - 0.72, advanced adenoma
AUC=0.68 - 0.77, carcinoma AUC=0.84 - 0.90; Figure S2). The OTUs that
were incorporated into the adenoma and advanced adenoma models largely
overlapped and those OTUs that were used to classify the carcinoma
samples were largely distinct from those of the other two models
{[}Figure 3A{]}. Among the OTUs that were shared across the three models
were those populations commonly considered as commensals (e.g.
\emph{Faecalibacterium}, \emph{Lachnospiraceae}, \emph{Bacteroides},
\emph{Dorea}, \emph{Anaerostipes}, and \emph{Roseburia}) {[}Figures
3B{]}. Although many of these OTUs were also included in the model
differentiating between patients with normal colons and those with
carcinoma, this model also included OTUs affiliated with populations
that have previously been associated with carcinoma
(\emph{Fusobacterium}, \emph{Porphyromonas}, \emph{Parvimonas})
{[}Figure S3{]}. Finally, we applied these three models to the pre and
post-treatment samples for each diagnosis group and quantified the
change in the positive probability of the model. A decrease in the
positive probability would indicate that the colon more closely
resembled that of a patient with a normal colon. There was no
significant change in the positive probability for the adenoma or
advanced adenoma groups {[}Figure 4{]}. The positive probability for the
pre and post-treatment samples from patients diagnosed with carcinoma
significantly decreased with treatment. Only, 6 of the 26 patients
(23.08\%) who were diagnosed with a carcinoma had a higher positive
probability after treatment; one of those was re-diagnosed with
carcinoma on the follow up visit. These results indicate that, although
there were changes in the microbiota associated with treatment, those
experienced by patients with carcinoma after treatment yielded gut
bacterial communities of greater similarity to that of a normal colon.

\textbf{\emph{Difficult to identify effects of specific treatments on
the change in the microbiota.}} The type of treatment that the patients
received varied across diagnosis groups. Those with adenomas and
advanced adenomas received surgical resection (adenoma, N=4; advanced
adenoma, N=4) or polyp removal during colonoscopy (adenoma, N=18;
advanced adenoma, N=15) and those with carcinomas received surgical
resection (N=12), surgical resection with chemotherapy (N=9), and
surgical resection with chemotherapy and radiation (N=5). We focused on
the patients with carcinoma and pooled those patients that received
chemotherapy with those that received chemotherapy and radiation to
improve our statistical power. We did not observe a significant
difference in the effect of these treatments on the number of observed
OTUs, Shannon diversity, or Shannon evenness (P-value \textgreater{}
0.05). Furthermore, there was not a significant difference in the effect
of the treatments on the amount of change in the community structure
(P-value = 0.298). Finally, the change in the positive probability was
not significantly different between the two treatment groups (P-value =
0.999). Due to the relatively small number of samples in each treatment
group, it was difficult to make a definitive statement regarding the
specific type of treatment on the amount of change in the structure of
the microbiota.

\newpage

\subsection{Discussion}\label{discussion}

This study builds upon previous work that demonstrated a role for the
microbiota in tumorigenesis and that microbiome-derived biomarkers can
detect colonic lesions {[}6--10,24--28{]}. Our study focused on
comparing the microbiota of patients diagnosed with adenoma, advanced
adenoma, and carcinoma before and after treatment. For all three groups
of patients, we observed changes in their microbiota. After treatment,
the microbiota of patients with carcinoma changed significantly more
than the other groups. This change resulted in communities that more
closely resembled those of patients with a normal colon. This may
suggest that treatment for carcinoma is not only successful for removing
the carcinoma but also at reducing the associated bacterial communities.
Understanding the effect of treatment on the microbiota of those
diagnosed with carcinomas may have important implications for reducing
disease recurrence. It is intriguing that it may be possible to use
microbiome-based biomarkers to not only predict the presence of lesions
but to assess the risk of recurrence.

Patients diagnosed with adenoma and advanced adenoma, however, did not
experience a shift towards a community structure that resembled those
with normal colons. This may be due to the fundamental differences
between the features of adenomas and advanced adenomas and carcinoma.
Specifically, carcinomas may create an inflammatory milieu that would
impact the structure of the community and removal of that stimulus would
alter said structure. In addition, it is possible that the difference
between the microbiota of patients with adenoma and advanced adenoma and
those with normal colons is subtle. This is supported by the reduced
ability of our models to correctly classify patients with adenomas and
advanced adenomas relative to those diagnosed with carcinomas {[}Figure
S2{]}. Given the irregular distribution of microbiota across patients in
the different diagnosis groups, it is possible that we lacked the
statistical power to adequately characterize the change in the
communities following treatment.

There was a subset of patients (6 of the 26 with carcinomas) who
demonstrated an elevated probability of carcinoma after treatment. This
may reflect an elevated risk of recurrence. The 23.08\% prevalence of
increased carcinoma probability from our study is within the expected
rate of recurrence (20-30\% {[}3,4{]}). We hypothesized that these
individuals may have had more severe tumors; however, the tumor severity
of these 6 individuals (3 with Stage II and 3 with Stage III) was
similar to the distribution observed among the other 20 patients. We
also hypothesized that we may have sampled these patients later than the
rest and their communities may have reverted to a carcinoma-associated
state; however, there was not a statistically significant difference in
the length of time between sample collection among those whose
probabilities increased or decreased (Wilcoxon Test; P-value = 0.56).
Finally, it is possible that these patients may not have responded to
treatment as well as the other 20 patients diagnosed with carcinoma and
so the microbiota may not have been impacted the same way. Again,
further studies looking at the role of the microbiota in recurrence are
needed to understand the dynamics following treatment.

Our final hypothesis was that the specific type of treatment altered the
structure of the microbiome. The treatment to remove adenomas and
advanced adenomas was either polyp removal or surgical resection whereas
it was surgical resection alone or in combination with chemotherapy or
with chemotherapy and radiation for individuals with carcinoma. Because
chemotherapy and radiation target rapidly growing cells, these
treatments would be more likely to cause a turnover of the colonic
epithelium driving a more significant change in the structure of the
microbiota. Although, we were able to test for an effect across these
specific types of treatment, the number of patients in each treatment
group was relatively small.

This study expands upon existing research that has established a role
for the microbiota in tumorigenesis and that demonstrated the utility of
microbiome-based biomarkers to predict the presence of colonic lesions.
The most exciting future direction from the current study is the
possibility that markers within the microbiota could be used to evaluate
the effect of treatment and predict recurrence for those diagnosed with
carcinoma. If such an approach is effective, it might be possible to
target the microbiota as part of adjuvant therapy. Our data provides
additional evidence on the importance of the microbiota in tumorigenesis
by addressing the recovery of the microbiota after treatment and opens
interesting avenues of research into how these changes may affect
recurrence.

\newpage

\subsection{Methods}\label{methods}

\textbf{\emph{Study Design and Patient Sampling.}} Sampling and design
have been previously reported in Baxter, et al {[}24{]}. Briefly, study
exclusion involved those who had already undergone surgery, radiation,
or chemotherapy, had colorectal cancer before a baseline fecal sample
could be obtained, had IBD, a known hereditary non-polyposis colorectal
cancer, or familial adenomatous polyposis. Samples used to build the
models for prediction were collected either prior to a colonoscopy or
between one and two weeks after initial colonoscopy. The bacterial
community has been shown to normalize back to a pre-colonoscopy
community within this time period {[}29{]}. Our study cohort consisted
of 67 individuals with an initial sample as described and a follow up
sample obtained between 188 - 546 days after treatment of lesion
{[}Table 1{]}. Patients were diagnosed by colonoscopic examination and
histopathological review of biopsies. Patients were classified as having
advanced adenoma if they had an adenoma greater than 1 cm, more than
three adenomas of any size, or an adenoma with villous histology. This
study was approved by the University of Michigan Institutional Review
Board. All study participants provided informed consent and the study
itself conformed to the guidelines set out by the Helsinki Declaration.

\textbf{\emph{16S rRNA Gene Sequencing.}} Sequencing was completed as
described by Kozich, et al. {[}30{]}. DNA extraction used the 96-well
Soil DNA isolation kit (MO BIO Laboratories) and an epMotion 5075
automated pipetting system (Eppendorf). The V4 variable region was
amplified and the resulting product was split between three sequencing
runs with normal, adenoma, and carcinoma evenly represented on each run.
Each group was randomly assigned to avoid biases based on sample
collection location. The pre and post-treatment samples were sequenced
on the same run.

\textbf{\emph{Sequence Processing.}} The mothur software package
(v1.37.5) was used to process the 16S rRNA gene sequences and has been
previously described {[}30{]}. The general workflow using mothur
included merging paired-end reads into contigs, filtering for low
quality contigs, aligning to the SILVA database {[}31{]}, screening for
chimeras using UCHIME {[}32{]}, classifying with a naive Bayesian
classifier using the Ribosomal Database Project (RDP){[}33{]}, and
clustered into Operational Taxonomic Units (OTUs) using a 97\%
similarity cutoff with an average neighbor clustering algorithm
{[}34{]}. The number of sequences for each sample was rarefied to 10523
to minimize the impacts of uneven sampling.

\textbf{\emph{Model Building.}} The Random Forest {[}35{]} algorithm was
used to create the model used to create the three models used. The
adenoma model classified normal versus adenoma, the advanced adenoma
model classified normal versus advanced adenoma, and the carcinoma model
classified normal versus carcinoma. The total number of individuals in
the treatment models was 67 individuals. There were a total of 22
individuals in the treatment adenoma model, 19 individuals in the
treatment advanced adenoma model, and 26 individuals in the treatment
carcinoma model {[}Table 1{]}. The total number of individuals in the
normal versus diagnosis models was 423 individuals {[}Table 2{]}. There
were a total of 239 individuals in the adenoma model, 262 individuals in
the advanced adenoma model, and 266 individuals in the carcinoma model
{[}Table 2{]}. Each model was then applied to our 67-person cohort
{[}Table 1{]} which assesed the prediction of pre-treatment adenoma
(adenoma n = 22) versus post-treatment adenoma (adenoma n = 0),
pre-treatment advanced adenoma (advanced adenoma n = 19) versus
post-treatment advanced adenoma (advanced adenoma n = 0), pre-treatment
carcinoma (carinoma n = 26) versus post-treatment carcinoma (carcinoma n
= 1).

The model included only OTU data obtained from 16S rRNA sequencing.
Non-binary data was checked for near zero variance and OTUs that had
near zero variance were removed. This pre-processing was performed with
the R package caret (v6.0.73). Optimization of the mtry hyper-parameter
involved making 100 different 80/20 (train/test) splits of the data
where normal and adenoma, normal and advanced adenoma, or normal and
carcinoma were represented in the same proportion within both the whole
data set and the 80/20 split. For each of the different splits, 20
repeated 10-fold cross validation was performed on the 80\% component to
optimize the mtry hyper-parameter by maximizing the AUC (Area Under the
Curve of the Receiver Operator Characteristic). The resulting model was
then tested on the hold out data obtained from the 20\% component. All
three models had an optimized mtry of 2. The hyper-parameter, mtry,
defines the number of variables to investigate at each split before a
new division of the data was created with the Random Forest model.

Assessment of the most important OTUs to the model involved counting the
number of times an OTU was present in the top 10\% of mean decrease in
accuracy (MDA) for each of the 100 different splits run. This was then
followed with filtering of this list to variables that were only present
in more than 50\% of these 100 runs. The final collated list of
variables was then run through the mtry optimization again. Once the
ideal mtry was found the entire sample set specific to normal versus
adenoma, normal versus advanced adenoma, or normal versus carcinoma was
used to create the final Random Forest model on which classifications on
the 67-person cohort was completed. For all three models the final
optimized mtry was 2. The only difference other than the sample set used
between treatment models and normal versus diagnosis models was that
only the top 10 OTUs were used to build each respective treatment model
by diagnosis group to help avoid model overfitting.T

\textbf{\emph{Statistical Analysis.}} The R software package (v3.3.2)
was used for all statistical analysis. Comparisons between bacterial
community structure utilized PERMANOVA {[}36{]} in the vegan package
(v2.4.1). Comparisons between probabilities as well as overall OTU
differences between pre and post-treatment samples utilized a paired
Wilcoxon ranked sum test. Where multiple comparison testing was
appropriate, a Benjamini-Hochberg (BH) correction was applied {[}37{]}
and a corrected P-value of less than 0.05 was considered significant.
Unless otherwise stated the P-values reported are those that were BH
corrected. Model rank importance was determined by obtaining the median
MDA from the 100, 20 repeated 10-fold cross validation and then ranking
from largest to smallest MDA.

\textbf{\emph{Reproducible Methods.}} A detailed and reproducible
description of how the data were processed and analyzed can be found at
\url{https://github.com/SchlossLab/Sze_followUps_2017}. Raw sequences
have been deposited into the NCBI Sequence Read Archive (SRP062005 and
SRP096978) and the necessary metadata can be found at
\url{https://www.ncbi.nlm.nih.gov/Traces/study/} and searching the
respective SRA study accession.

\newpage

\textbf{Figure 1: General Differences between Adenoma, Advanced Adenoma,
and Carcinoma Groups After Treatment.} A) A significant difference was
found between the adenoma and carcinoma group for thetayc (P-value =
5.36e-05). Solid black points represent the median value for each
diagnosis group. B) NMDS of the pre and post-treatment samples for the
adenoma group. C) NMDS of the pre and post-treatment samples for the
advanced adenoma group. D) NMDS of the pre and post-treatment samples
for the carcinoma group.

\textbf{Figure 2: The 10 OTUs used to classify treatment for Adenoma,
Advanced Adenoma, and Carcinoma.} A) Adenoma OTUs. B) Advanced Adenoma
OTUs. C) Carcinoma OTUs.

\textbf{Figure 3: Treatment Response Based on Models Built for Adenoma,
Advanced Adenoma, or Carcinoma.} A) Positive probability change from
initial to follow up sample in those with adenoma. B) Positive
probability change from initial to follow up sample in those with
advanced adenoma. C) Positive probability change from initial to follow
up sample in those with carcinoma.

\textbf{Figure 4: Common OTUs to All Models.} A) Venn diagram showing
the OTU overlap between each model. B) For each common OTU the lowest
taxonomic identification and importance rank for each model run is
shown.

\newpage

\textbf{Table 1: Demographic Data of pre and Post Treatment Cohort}

\begin{longtable}[]{@{}lccc@{}}
\toprule
& Adenoma & Advanced Adenoma & Carcinoma\tabularnewline
\midrule
\endhead
n & 22 & 19 & 26\tabularnewline
Age (Mean ± SD) & 61.68 ± 7.2 & 63.11 ± 10.9 & 61.65 ±
12.9\tabularnewline
Sex (\%F) & 36.36 & 36.84 & 42.31\tabularnewline
BMI (Mean ± SD) & 26.86 ± 3.9 & 25.80 ± 4.7 & 28.63 ± 7.2\tabularnewline
Caucasian (\%) & 95.45 & 84.21 & 96.15\tabularnewline
\bottomrule
\end{longtable}

\newpage

\textbf{Table 2: Demographic Data of Training Cohort}

\begin{longtable}[]{@{}lcccc@{}}
\toprule
& Normal & Adenoma & Advanced Adenoma & Carcinoma\tabularnewline
\midrule
\endhead
n & 172 & 67 & 90 & 94\tabularnewline
Age (Mean ± SD) & 54.29 ± 9.9 & 63.01 ± 13.1 & 64.07 ± 11.3 & 64.37 ±
12.9\tabularnewline
Sex (\%F) & 64.53 & 46.27 & 37.78 & 43.62\tabularnewline
BMI (Mean ± SD) & 26.96 ± 5.3 & 25.68 ± 4.8 & 26.66 ± 4.9 & 29.27 ±
6.7\tabularnewline
Caucasian (\%) & 87.79 & 92.54 & 92.22 & 94.68\tabularnewline
\bottomrule
\end{longtable}

\newpage

\textbf{Figure S1: Distribution of P-values from Paired Wilcoxson
Analysis of All OTUs Before and After Treatment}

\textbf{Figure S2: ROC Curves of the Adenoma, Advanced Adenoma, and
Carcinoma Models.} A) Adenoma ROC curve: The light green shaded areas
represent the range of values of a 100 different 80/20 splits of the
test set data and the dark green line represents the model using 100\%
of the data set and what was used for subsequent classification. B)
Advanced Adenoma ROC curve: The light yellow shaded areas represent the
range of values of a 100 different 80/20 splits of the test set data and
the dark yellow line represents the model using 100\% of the data set
and what was used for subsequent classification. C) Carcinoma ROC curve:
The light red shaded areas represent the range of values of a 100
different 80/20 splits of the test set data and the dark red line
represents the model using 100\% of the data set and what was used for
subsequent classification.

\textbf{Figure S3: Summary of Important OTUs for the Adenoma, Advanced
Adenoma, and Carcinoma Models.} A) MDA of the most important variables
in the adenoma model. The dark green point represents the mean and the
lighter green points are the value of each of the 100 different runs. B)
Summary of Important Variables in the advanced adenoma model. MDA of the
most important variables in the SRN model. The dark yellow point
represents the mean and the lighter yellow points are the value of each
of the 100 different runs. C) MDA of the most important variables in the
carcinoma model. The dark red point represents the mean and the lighter
red points are the value of each of the 100 different runs.

\newpage

\subsection{Declarations}\label{declarations}

\subsubsection{Ethics approval and consent to
participate}\label{ethics-approval-and-consent-to-participate}

The University of Michigan Institutional Review Board approved this
study, and all subjects provided informed consent. This study conformed
to the guidelines of the Helsinki Declaration.

\subsubsection{Consent for publication}\label{consent-for-publication}

Not applicable.

\subsubsection{Availability of data and
material}\label{availability-of-data-and-material}

A detailed and reproducible description of how the data were processed
and analyzed can be found at
\url{https://github.com/SchlossLab/Sze_followUps_2017}. Raw sequences
have been deposited into the NCBI Sequence Read Archive (SRP062005 and
SRP096978) and the necessary metadata can be found at
\url{https://www.ncbi.nlm.nih.gov/Traces/study/} and searching the
respective SRA study accession.

\subsubsection{Competing Interests}\label{competing-interests}

All authors declare that they do not have any relevant competing
interests to report.

\subsubsection{Funding}\label{funding}

This study was supported by funding from the National Institutes of
Health to P. Schloss (R01GM099514, P30DK034933) and to the Early
Detection Research Network (U01CA86400).

\subsubsection{Authors' contributions}\label{authors-contributions}

All authors were involved in the conception and design of the study. MAS
analyzed the data. NTB processed samples and analyzed the data. All
authors interpreted the data. MAS and PDS wrote the manuscript. All
authors reviewed and revised the manuscript. All authors read and
approved the final manuscript.

\subsubsection{Acknowledgements}\label{acknowledgements}

The authors thank the Great Lakes-New England Early Detection Research
Network for providing the fecal samples that were used in this study. We
would also like to thank Amanda Elmore for reviewing and correcting code
error and providing feedback on manuscript drafts. We would also like to
thank Nicholas Lesniak for providing feedback on manuscript drafts.

\newpage

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\hypertarget{ref-siegel_cancer_2016}{}
1. Siegel RL, Miller KD, Jemal A. Cancer statistics, 2016. CA: a cancer
journal for clinicians. 2016;66:7--30.

\hypertarget{ref-haggar_colorectal_2009}{}
2. Haggar FA, Boushey RP. Colorectal cancer epidemiology: Incidence,
mortality, survival, and risk factors. Clinics in Colon and Rectal
Surgery. 2009;22:191--7.

\hypertarget{ref-hellinger_reoperation_2006}{}
3. Hellinger MD, Santiago CA. Reoperation for recurrent colorectal
cancer. Clinics in Colon and Rectal Surgery. 2006;19:228--36.

\hypertarget{ref-ryuk_predictive_2014}{}
4. Ryuk JP, Choi G-S, Park JS, Kim HJ, Park SY, Yoon GS, et al.
Predictive factors and the prognosis of recurrence of colorectal cancer
within 2 years after curative resection. Annals of Surgical Treatment
and Research. 2014;86:143--51.

\hypertarget{ref-national_cancer_institute_seer_nodate}{}
5. Institute NC. SEER Cancer Stat Facts: Colon and Rectum Cancer
{[}Internet{]}. {[}cited 2017 Apr 27{]}. Available from:
\url{http://seer.cancer.gov/statfacts/html/colorect.html}

\hypertarget{ref-goodwin_polyamine_2011}{}
6. Goodwin AC, Destefano Shields CE, Wu S, Huso DL, Wu X, Murray-Stewart
TR, et al. Polyamine catabolism contributes to enterotoxigenic
Bacteroides fragilis-induced colon tumorigenesis. Proceedings of the
National Academy of Sciences of the United States of America.
2011;108:15354--9.

\hypertarget{ref-abed_fap2_2016}{}
7. Abed J, Emgård JEM, Zamir G, Faroja M, Almogy G, Grenov A, et al.
Fap2 Mediates Fusobacterium nucleatum Colorectal Adenocarcinoma
Enrichment by Binding to Tumor-Expressed Gal-GalNAc. Cell Host \&
Microbe. 2016;20:215--25.

\hypertarget{ref-arthur_microbial_2014}{}
8. Arthur JC, Gharaibeh RZ, Mühlbauer M, Perez-Chanona E, Uronis JM,
McCafferty J, et al. Microbial genomic analysis reveals the essential
role of inflammation in bacteria-induced colorectal cancer. Nature
Communications. 2014;5:4724.

\hypertarget{ref-kostic_fusobacterium_2013}{}
9. Kostic AD, Chun E, Robertson L, Glickman JN, Gallini CA, Michaud M,
et al. Fusobacterium nucleatum potentiates intestinal tumorigenesis and
modulates the tumor-immune microenvironment. Cell Host \& Microbe.
2013;14:207--15.

\hypertarget{ref-wu_human_2009}{}
10. Wu S, Rhee K-J, Albesiano E, Rabizadeh S, Wu X, Yen H-R, et al. A
human colonic commensal promotes colon tumorigenesis via activation of T
helper type 17 T cell responses. Nature Medicine. 2009;15:1016--22.

\hypertarget{ref-zackular_manipulation_2016}{}
11. Zackular JP, Baxter NT, Chen GY, Schloss PD. Manipulation of the Gut
Microbiota Reveals Role in Colon Tumorigenesis. mSphere. 2016;1.

\hypertarget{ref-zackular_gut_2013}{}
12. Zackular JP, Baxter NT, Iverson KD, Sadler WD, Petrosino JF, Chen
GY, et al. The gut microbiome modulates colon tumorigenesis. mBio.
2013;4:e00692--00613.

\hypertarget{ref-baxter_structure_2014}{}
13. Baxter NT, Zackular JP, Chen GY, Schloss PD. Structure of the gut
microbiome following colonization with human feces determines colonic
tumor burden. Microbiome. 2014;2:20.

\hypertarget{ref-flynn_metabolic_2016}{}
14. Flynn KJ, Baxter NT, Schloss PD. Metabolic and Community Synergy of
Oral Bacteria in Colorectal Cancer. mSphere. 2016;1.

\hypertarget{ref-wang_structural_2012}{}
15. Wang T, Cai G, Qiu Y, Fei N, Zhang M, Pang X, et al. Structural
segregation of gut microbiota between colorectal cancer patients and
healthy volunteers. The ISME journal. 2012;6:320--9.

\hypertarget{ref-chen_decreased_2013}{}
16. Chen H-M, Yu Y-N, Wang J-L, Lin Y-W, Kong X, Yang C-Q, et al.
Decreased dietary fiber intake and structural alteration of gut
microbiota in patients with advanced colorectal adenoma. The American
Journal of Clinical Nutrition. 2013;97:1044--52.

\hypertarget{ref-chen_human_2012}{}
17. Chen W, Liu F, Ling Z, Tong X, Xiang C. Human intestinal lumen and
mucosa-associated microbiota in patients with colorectal cancer. PloS
One. 2012;7:e39743.

\hypertarget{ref-shen_molecular_2010}{}
18. Shen XJ, Rawls JF, Randall T, Burcal L, Mpande CN, Jenkins N, et al.
Molecular characterization of mucosal adherent bacteria and associations
with colorectal adenomas. Gut Microbes. 2010;1:138--47.

\hypertarget{ref-kostic_genomic_2012}{}
19. Kostic AD, Gevers D, Pedamallu CS, Michaud M, Duke F, Earl AM, et
al. Genomic analysis identifies association of Fusobacterium with
colorectal carcinoma. Genome Research. 2012;22:292--8.

\hypertarget{ref-feng_gut_2015}{}
20. Feng Q, Liang S, Jia H, Stadlmayr A, Tang L, Lan Z, et al. Gut
microbiome development along the colorectal adenoma-carcinoma sequence.
Nature Communications. 2015;6:6528.

\hypertarget{ref-dejea_microbiota_2014}{}
21. Dejea CM, Wick EC, Hechenbleikner EM, White JR, Mark Welch JL,
Rossetti BJ, et al. Microbiota organization is a distinct feature of
proximal colorectal cancers. Proceedings of the National Academy of
Sciences of the United States of America. 2014;111:18321--6.

\hypertarget{ref-mima_fusobacterium_2015}{}
22. Mima K, Sukawa Y, Nishihara R, Qian ZR, Yamauchi M, Inamura K, et
al. Fusobacterium nucleatum and T Cells in Colorectal Carcinoma. JAMA
oncology. 2015;1:653--61.

\hypertarget{ref-arthur_intestinal_2012}{}
23. Arthur JC, Perez-Chanona E, Mühlbauer M, Tomkovich S, Uronis JM, Fan
T-J, et al. Intestinal inflammation targets cancer-inducing activity of
the microbiota. Science (New York, N.Y.). 2012;338:120--3.

\hypertarget{ref-baxter_microbiota-based_2016}{}
24. Baxter NT, Ruffin MT, Rogers MAM, Schloss PD. Microbiota-based model
improves the sensitivity of fecal immunochemical test for detecting
colonic lesions. Genome Medicine. 2016;8:37.

\hypertarget{ref-zeller_potential_2014}{}
25. Zeller G, Tap J, Voigt AY, Sunagawa S, Kultima JR, Costea PI, et al.
Potential of fecal microbiota for early-stage detection of colorectal
cancer. Molecular Systems Biology. 2014;10:766.

\hypertarget{ref-zackular_human_2014}{}
26. Zackular JP, Rogers MAM, Ruffin MT, Schloss PD. The human gut
microbiome as a screening tool for colorectal cancer. Cancer Prevention
Research (Philadelphia, Pa.). 2014;7:1112--21.

\hypertarget{ref-yu_metagenomic_2017}{}
27. Yu J, Feng Q, Wong SH, Zhang D, Liang QY, Qin Y, et al. Metagenomic
analysis of faecal microbiome as a tool towards targeted non-invasive
biomarkers for colorectal cancer. Gut. 2017;66:70--8.

\hypertarget{ref-warren_co-occurrence_2013}{}
28. Warren RL, Freeman DJ, Pleasance S, Watson P, Moore RA, Cochrane K,
et al. Co-occurrence of anaerobic bacteria in colorectal carcinomas.
Microbiome. 2013;1:16.

\hypertarget{ref-obrien_impact_2013}{}
29. O'Brien CL, Allison GE, Grimpen F, Pavli P. Impact of colonoscopy
bowel preparation on intestinal microbiota. PloS One. 2013;8:e62815.

\hypertarget{ref-kozich_development_2013}{}
30. Kozich JJ, Westcott SL, Baxter NT, Highlander SK, Schloss PD.
Development of a dual-index sequencing strategy and curation pipeline
for analyzing amplicon sequence data on the MiSeq Illumina sequencing
platform. Applied and Environmental Microbiology. 2013;79:5112--20.

\hypertarget{ref-pruesse_silva_2007}{}
31. Pruesse E, Quast C, Knittel K, Fuchs BM, Ludwig W, Peplies J, et al.
SILVA: A comprehensive online resource for quality checked and aligned
ribosomal RNA sequence data compatible with ARB. Nucleic Acids Research.
2007;35:7188--96.

\hypertarget{ref-edgar_uchime_2011}{}
32. Edgar RC, Haas BJ, Clemente JC, Quince C, Knight R. UCHIME improves
sensitivity and speed of chimera detection. Bioinformatics (Oxford,
England). 2011;27:2194--200.

\hypertarget{ref-wang_naive_2007}{}
33. Wang Q, Garrity GM, Tiedje JM, Cole JR. Naive Bayesian classifier
for rapid assignment of rRNA sequences into the new bacterial taxonomy.
Applied and Environmental Microbiology. 2007;73:5261--7.

\hypertarget{ref-schloss_assessing_2011}{}
34. Schloss PD, Westcott SL. Assessing and improving methods used in
operational taxonomic unit-based approaches for 16S rRNA gene sequence
analysis. Applied and Environmental Microbiology. 2011;77:3219--26.

\hypertarget{ref-breiman_random_2001}{}
35. Breiman L. Random Forests. Machine Learning {[}Internet{]}. 2001
{[}cited 2013 Feb 7{]};45:5--32. Available from:
\href{http://link.springer.com/article/10.1023/A\%3A1010933404324\%20http://link.springer.com/article/10.1023\%2FA\%3A1010933404324?LI=true}{http://link.springer.com/article/10.1023/A\%3A1010933404324 http://link.springer.com/article/10.1023\%2FA\%3A1010933404324?LI=true}

\hypertarget{ref-anderson_permanova_2013}{}
36. Anderson MJ, Walsh DCI. PERMANOVA, ANOSIM, and the Mantel test in
the face of heterogeneous dispersions: What null hypothesis are you
testing? Ecological Monographs {[}Internet{]}. 2013 {[}cited 2017 Jan
5{]};83:557--74. Available from:
\url{http://doi.wiley.com/10.1890/12-2010.1}

\hypertarget{ref-benjamini_controlling_1995}{}
37. Benjamini Y, Hochberg Y. Controlling the false discovery rate: A
practical and powerful approach to multiple testing. Journal of the
Royal Statistical Society. Series B (Methodological). 1995;57:289--300.

\end{document}