### 2017-09-13 submission/manuscript_R2.tex

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}
\usepackage{microtype}
\usepackage{pdfpages}
\usepackage{comment}
\usepackage{lscape}

\begin{document}

\section{Normalization of the microbiota in patients after treatment for
colonic
lesions}\label{normalization-of-the-microbiota-in-patients-after-treatment-for-colonic-lesions}

\begin{center}
\vspace{25mm}

Marc A Sze${^1}$, Nielson T Baxter${^1}$${^,}$${^2}$, Mack T Ruffin IV${^3}$, Mary AM Rogers${^2}$, and Patrick D Schloss${^1}$${^\dagger}$

\vspace{20mm}

$\dagger$ To whom correspondence should be addressed: pschloss@umich.edu

$1$ Department of Microbiology and Immunology, University of Michigan, Ann Arbor, MI

$2$ Department of Internal Medicine, University of Michigan, Ann Arbor, MI

$3$ Department of Family Medicine and Community Medicine, Penn State Hershey Medical Center, Hershey, PA

\end{center}

Co-author e-mails:

\begin{itemize}
\tightlist
\item
  \href{mailto:marcsze@med.umich.med.edu}{\nolinkurl{marcsze@med.umich.med.edu}}
\item
  \href{mailto:ntbaxter@umich.edu}{\nolinkurl{ntbaxter@umich.edu}}
\item
  \href{mailto:mruffin@pennstatehealth.psu.edu}{\nolinkurl{mruffin@pennstatehealth.psu.edu}}
\item
  \href{mailto:maryroge@med.umich.edu}{\nolinkurl{maryroge@med.umich.edu}}
\end{itemize}

\newpage

\linenumbers

\subsection{Abstract}\label{abstract}

\textbf{Background.} Despite growing evidence that members of the gut
microbiota can drive tumorigenesis, little is known about what happens
to it after treatment for an adenoma or carcinoma. This study tests
whether treatment alters the abundance of bacterial populations
associated with disease versus those with a normal colon by sequencing
the 16S rRNA genes in feces of 67 individuals before and after treatment
(adenoma n = 22), advanced adenoma n = 19), and carcinoma n = 26).

\textbf{Results.} The communities from patients with carcinomas changed
significantly more than those with adenoma following treatment (P-value
\textless{} 0.001). Although Random Forest models classified pre and
post-treatment samples for each diagnosis group, there was little
overlap between the OTUs that were indicative of each treatment.
Training Random Forest models on a larger cohort that contained
individuals with normal colons, adenomas, advanced adenomas, and
carcinomas we assessed the positive probability of having one of three
diagnoses and whether the pre and post-treatment samples were classified
the same. Patients who had carcinomas changed towards a microbial milieu
that resembles the normal colon after treatment (P-value \textless{}
0.05). There was no significant difference in the microbiota of
individuals treated with surgery alone and those treated with
chemotherapy or chemotherapy and radiation (P-value \textgreater{}
0.05).

\textbf{Conclusions.} It was difficult to identify significant
differences between pre and post-treatment samples from patients with
adenoma and advanced adenoma. However, for carcinomas, not only were
there large changes in pre versus post-treatment samples, but these
changes were towards a more normal microbiota.

\newpage

\subsection{Importance}\label{importance}

This study helps to clarify what happens to the microbiota after
generalized treatment for an adenoma or carcinoma. It is a necessary
first step to ultimately investigating how the microbiota changes in
those with carcinoma recurrence. This is important since a large number
of those treated for carcinoma will have a recurrence. Finding better
ways to stratify those most and least likely to recur using the
microbiota could provide a cheaper and more accurate alternative to the
current status quo. Our results suggest that the microbiota could be
used to stratify individuals for risk of recurrence and provides a
starting point for future studies into this area.

\newpage

\subsection{Background}\label{background}

Colorectal cancer (CRC) is the third most common cause of cancer deaths
in the United States (1, 2). Disease mortality has significantly
decreased, predominately due to improvements in screening (2). Despite
these improvements, there are still approximately 50,000 CRC-related
deaths per year in the United States (1). Current estimates indicate
that 20-30\% of those who undergo treatment will experience recurrence
and 35\% of all patients will die within five years (3--5).
Identification of methods to assess patients' risk of recurrence is of
great importance to reduce mortality and healthcare costs.

There is growing evidence that the gut microbiota is involved in the
progression of CRC. Mouse-based studies have identified populations of
\emph{Bacteroides fragilis}, \emph{Escherichia coli}, and
\emph{Fusobacterium nucleatum} that alter disease progression (6--10).
Furthermore, studies that shift the structure of the microbiota through
the use of antibiotics or inoculation of germ free mice with human feces
have shown that varying community compositions can result in varied
tumor burden (11--13). Collectively, these studies support the
hypothesis that the microbiota can alter the amount of inflammation in
the colon and with it the rate of tumorigenesis (14).

Building upon this evidence, several human studies have identified
unique signatures of colonic lesions (15--20). One line of research has
identified community-level differences between those bacteria that are
found on and adjacent to colonic lesions and have supported a role for
\emph{Bacteroides fragilis}, \emph{Escherichia coli}, and
\emph{Fusobacterium nucleatum} in tumorigenesis (21--23). Others have
proposed feces-based biomarkers that could be used to diagnose the
presence of colonic adenomas and carcinomas (24--26). These studies have
associated \emph{Fusobacterium nucleatum} and other oral pathogens with
colonic lesions (adenoma, advanced adenoma, and carcinoma). They have
also noted that the loss of bacteria generally thought to produce short
chain fatty acids, which can suppress inflammation, is associated with
colonic lesions. This suggests that gut bacteria have a role in
tumorigenesis with potential as useful biomarkers for aiding in the
early detection of disease (21--26).

Despite advances in understanding the role between the gut microbiota
and colonic tumorigenesis, we still do not understand how treatments
including resection, chemotherapy, and/or radiation affect the
composition of the gut microbiota. If the microbial community drives
tumorigenesis then one would hypothesize that treatment to remove a
lesion would affect the microbiota and risk of recurrence. To test this
hypothesis, we addressed two related questions: Does treatment affect
the colonic microbiota in a predictable manner? If so, does the
treatment alter the community to more closely resemble that of
individuals with normal colons?

We answered these questions by sequencing the V4 region of 16S rRNA
genes amplified from fecal samples of individuals with adenoma, advanced
adenoma, and carcinomas pre and post-treatment. We used classical
community analysis to compare the alpha and beta-diversity of
communities pre and post-treatment. Next, we generated Random Forest
models to identify bacterial populations that were indicative of
treatment for each diagnosis group. Finally, we measured the predictive
probabilities to assess whether treatment yielded bacterial communities
similar to those individuals with normal colons. We found that treatment
alters the composition of the gut microbiota and that, for those with
carcinomas, the gut microbiota shifted more towards that of a normal
colon after treatment. In the individuals with carcinomas, no difference
was found by the type of treatment (surgery alone, surgery with
chemotherapy, surgery with chemotherapy and radiation). Understanding
how the community responds to these treatments could be a valuable tool
for identifying biomarkers to quantify the risk of recurrence and the
likelihood of survival.

\newpage

\subsection{Results}\label{results}

\textbf{\emph{Treatment for colonic lesions alters the bacterial
community structure}} Within our 67-person cohort we tested whether the
microbiota of patients with adenoma (N = 22), advanced adenoma (N = 19),
or carcinoma (N = 26) had any broad differences between pre and
post-treatment samples {[}Table 1{]}. None of the individuals in this
study had any recorded antibiotic usage that was not associated with
surgical treatment of their respective lesion. The structure of the
microbial communities of the pre and post-treatment samples differed, as
measured by the \(\theta\)\textsubscript{YC} beta diversity metric
{[}Figure 1A{]}. We found that the communities obtained pre and
post-treatment among the patients with carcinomas changed significantly
more than those patients with adenoma (P-value \textless{} 0.001). There
were no significant differences in the amount of change observed between
the patients with adenoma and advanced adenoma or between the patients
with advanced adenoma and carcinoma (P-value \textgreater{} 0.05). Next,
we tested whether there was a consistent direction in the change in the
community structure between the pre and post-treatment samples for each
of the diagnosis groups {[}Figure 1B-D{]}. We only observed a consistent
shift in community structure for the patients with carcinoma when using
a PERMANOVA test (adenoma P-value = 0.999, advanced adenoma P-value =
0.945, and carcinoma P-value = 0.005). Finally, we measured the number
of observed OTUs, Shannon evenness, and Shannon diversity in the pre and
post-treatment samples and did not observe a significant change for any
of the diagnosis groups (P-value \textgreater{} 0.05) {[}Table S1{]}.

\textbf{\emph{The treatment of lesions are not consistent across
diagnosis groups.}} We used two approaches to identify those bacterial
populations that change between the two samples for each diagnosis
group. First, we sought to identify individual OTUs that could account
for the change in overall community structure. However, using a paired
Wilcoxon test we were unable to identify any OTUs that were
significantly different in the pre and post-treatment groups (P-value
\textgreater{} 0.05). It is likely that high inter-individual variation
and the irregular distribution of OTUs across individuals limited the
statistical power of the test. We attempted to overcome these problems
by using Random Forest models to identify collections of OTUs that would
allow us to differentiate between pre and post-treatment samples from
each of the diagnosis groups. The models overall performed moderately,
with the carcinoma model being the best (adenoma AUC range = 0.54 -
0.83, advanced adenoma AUC range = 0.34 - 0.65, carcinoma AUC range =
0.82 - 0.98). Interestingly, the top 10 most important OTUs by MDA that
were used for each model had little overlap with each other {[}Figure
2{]}. Although treatment had an impact on the overall community
structure, the effect of treatment was not consistent across patients
and diagnosis groups. Both the adenoma and carcioma treatment models had
AUCs that were significantly higher than a random model permutation
(P-value \textless{} 0.0001).

\textbf{\emph{Post-treatment samples from patients with carcinoma more
closely resemble those of a normal colon.}} Next, we determined whether
treatment changed the microbiota in a way that the post-treatment
communities resembled that of patients with normal colons. To test this,
we used an expanded cohort of 423 individuals that were diagnosed under
the same protocol as having normal colons or colons with adenoma,
advanced adenoma, or carcinoma {[}Table 2{]}. We then constructed Random
Forest models to classify the study samples, with the 3 diagnosis groups
(adenoma, advanced adenoma, or carcinoma), or having a normal colon. The
models performed moderately with CRC being the best (adenoma AUC range =
0.50 - 0.62, advanced adenoma AUC range = 0.53 - 0.67, carcinoma AUC
range = 0.71 - 0.82; Figure S1). The OTUs that were in the top 10\% of
importance for the adenoma and advanced adenoma models largely
overlapped and those OTUs that were used to classify the carcinoma
samples were largely distinct from those of the other two models
{[}Figure 3A{]}. Among the OTUs that were shared across the three models
were those populations generally considered beneficial to their host
(e.g. \emph{Faecalibacterium}, \emph{Lachnospiraceae},
\emph{Bacteroides}, \emph{Dorea}, \emph{Anaerostipes}, and
\emph{Roseburia}) {[}Figures 3B{]}. Although many of important OTUs in
the top 10\% were also included in the model differentiating between
patients with normal colons and those with carcinoma, this model also
included OTUs affiliated with populations that have previously been
associated with carcinoma (\emph{Fusobacterium}, \emph{Porphyromonas},
\emph{Parvimonas}) (24--26) {[}Figure S2{]} with some individuals
showing a marked decrease in relative abundance {[}Figure S3{]}.
Finally, we applied these three models to the pre and post-treatment
samples for each diagnosis group and quantified the change in the
positive probability of the model. A decrease in the positive
probability would indicate that the microbiota more closely resembled
that of a patient with a normal colon. There was no significant change
in the positive probability for the adenoma or advanced adenoma groups
(P-value \textgreater{} 0.05) {[}Figure 4{]}. The positive probability
for the pre and post-treatment samples from patients diagnosed with
carcinoma significantly decreased with treatment, suggesting a shift
toward a normal microbiota for most individuals (P-value = 0.001). Only,
7 of the 26 patients (26.92\%) who were diagnosed with a carcinoma had a
higher positive probability after treatment; one of those was
re-diagnosed with carcinoma on the follow up visit. These results
indicate that, although there were changes in the microbiota associated
with treatment, those experienced by patients with carcinoma after
treatment yielded gut bacterial communities of greater similarity to
that of a normal colon.

\textbf{\emph{Difficult to identify effects of specific treatments on
the change in the microbiota.}} The type of treatment that the patients
received varied across diagnosis groups. Those with adenomas and
advanced adenomas received surgical resection (adenoma, N=4; advanced
adenoma, N=4) or polyp removal during colonoscopy (adenoma, N=18;
advanced adenoma, N=15) and those with carcinomas received surgical
resection (N=12), surgical resection with chemotherapy (N=9), and
surgical resection with chemotherapy and radiation (N=5). Regardless of
treatment used there was no significant difference in the effect of
these treatments on the number of observed OTUs, Shannon diversity, or
Shannon evenness (P-value \textgreater{} 0.05). Furthermore, there was
not a significant difference in the effect of the treatments on the
amount of change in the community structure (P-value = 0.375). Finally,
the change in the positive probability was not significantly different
between any of the treatment groups (P-value = 0.375). Due to the
relatively small number of samples in each treatment group, it was
difficult to make a definitive statement regarding the specific type of
treatment on the amount of change in the structure of the microbiota.

\newpage

\subsection{Discussion}\label{discussion}

Our study focused on comparing the microbiota of patients diagnosed with
adenoma, advanced adenoma, and carcinoma before and after treatment. For
all three groups of patients, we observed changes in their microbiota.
Some of these changes, specifically for adenoma, may be due to normal
temporal variation, however, those with advanced adenoma and carcinoma
clearly had large microbiota changes. After treatment, the microbiota of
patients with carcinoma changed significantly more than the other
groups. This change resulted in communities that more closely resembled
those of patients with a normal colon. This may suggest that treatment
for carcinoma is not only successful for removing the carcinoma but also
at reducing the associated bacterial communities. Understanding the
effect of treatment on the microbiota of those diagnosed with carcinomas
may have important implications for reducing disease recurrence. It is
intriguing that it may be possible to use microbiome-based biomarkers to
not only predict the presence of lesions but to also assess the risk of
recurrence due to these changes in the microbiota.

Patients diagnosed with adenoma and advanced adenoma, however, did not
experience a shift towards a community structure that resembled those
with normal colons. This may be due to the fundamental differences
between the features of adenomas and advanced adenomas and carcinoma.
Specifically, carcinomas may create an inflammatory milieu that would
impact the structure of the community and removal of that stimulus would
alter said structure. It is possible that the difference between the
microbiota of patients with adenoma and advanced adenoma and those with
normal colons is subtle. This is supported by the reduced ability of our
models to correctly classify patients with adenomas and advanced
adenomas relative to those diagnosed with carcinomas {[}Figure S1{]}.
Given the irregular distribution of microbiota across patients in the
different diagnosis groups, it is possible that we lacked the
statistical power to adequately characterize the change in the
communities following treatment.

There was a subset of patients (7 of the 26 with carcinomas) who
demonstrated an elevated probability of carcinoma after treatment. This
may reflect an elevated risk of recurrence. The 26.92\% prevalence of
increased carcinoma probability from our study is within the expected
rate of recurrence (20-30\% (3, 4)). We hypothesized that these
individuals may have had more severe tumors; however, the tumor severity
of these 7 individuals (1 with Stage I, 3 with Stage II, and 3 with
Stage III) was similar to the distribution observed among the other 19
patients. We also hypothesized that we may have sampled these patients
later than the rest and their communities may have reverted to a
carcinoma-associated state; however, there was not a statistically
significant difference in the length of time between sample collection
among those whose probabilities increased (331 (246 - 358) days) or
decreased (364 (301 - 434) days) (Wilcoxon Test; P-value = 0.39) (all
days data displayed as median (IQR)). Finally, it is possible that these
patients may not have responded to treatment as well as the other 19
patients diagnosed with carcinoma and so the microbiota may not have
been impacted the same way. Again, further studies looking at the role
of the microbiota in recurrence are needed to understand the dynamics
following treatment.

Our final hypothesis was that the specific type of treatment altered the
structure of the microbiome. The treatment to remove adenomas and
advanced adenomas was either polyp removal or surgical resection whereas
it was surgical resection alone or in combination with chemotherapy or
with chemotherapy and radiation for individuals with carcinoma. Because
chemotherapy and radiation target rapidly growing cells, these
treatments would be more likely to cause a turnover of the colonic
epithelium driving a more significant change in the structure of the
microbiota. Although, we were able to test for an effect across these
specific types of treatment, the number of patients in each treatment
group was relatively small. Finally, those undergoing surgery would have
received antibiotics and this may be a potential confounder. However,
our pre-treatment stool samples were obtained before the surgery and the
post-treatment samples were obtained long after any effects due to
antibiotic administration on the microbiome would be expected to occur
(344 (266 - 408) days). We also found no difference in the community
structure of those that received surgery and those that did not as a
treatment for adenoma or advanced adenoma.

\subsection{Conclusion}\label{conclusion}

This study expands upon existing research that has established a role
for the microbiota in tumorigenesis and that demonstrated the utility of
microbiome-based biomarkers to predict the presence of colonic lesions.
We were surprised by the lack of a consistent signal that was associated
with treatment of patients with adenomas or advanced adenomas. The lack
of a large effect size may be due to differences in the role of bacteria
in the formation of adenomas and carcinomas or it could be due to
differences in the behaviors and medications within these classes of
patients. One of the most exciting of these future directions is the
possibility that markers within the microbiota could be used to
potentially evaluate the effect of treatment and predict recurrence for
those diagnosed with carcinoma. If such an approach is effective, it
might be possible to target the microbiota as part of adjuvant therapy,
if the biomarkers identified play a key role in the disease process. Our
data provides additional evidence on the importance of the microbiota in
tumorigenesis by addressing the recovery of the microbiota after
treatment and opens interesting avenues of research into how these
changes may affect recurrence.

\newpage

\subsection{Methods}\label{methods}

\textbf{\emph{Study Design and Patient Sampling.}} Sampling and design
have been previously reported in Baxter, et al (24). Briefly, samples
were stored on ice for at least 24h before freezing. Although we cannot
exclude that this sampling protocol may have impacted the gut microbiota
composition all samples were subjected to the same methodology. Study
exclusion involved those who had already undergone surgery, radiation,
or chemotherapy, had colorectal cancer before a baseline fecal sample
could be obtained, had IBD, a known hereditary non-polyposis colorectal
cancer, or familial adenomatous polyposis. Samples used to build the
models for prediction were collected either prior to a colonoscopy or
between one and two weeks after initial colonoscopy. The bacterial
community has been shown to normalize back to a pre-colonoscopy
community within this time period (27). Our study cohort consisted of 67
individuals with an initial sample as described and a follow up sample
obtained between 188 - 546 days after treatment of lesion {[}Table 1{]}.
Patients were diagnosed by colonoscopic examination and
histopathological review of any biopsies taken. Patients were classified
as having advanced adenoma if they had an adenoma greater than 1 cm,
more than three adenomas of any size, or an adenoma with villous
histology. This study was approved by the University of Michigan
Institutional Review Board. All study participants provided informed
consent and the study itself conformed to the guidelines set out by the
Helsinki Declaration.

\textbf{\emph{Treatment.}} For this study treatment refers specifically
to the removal of a lesion with or without chemotherapy and radiation.
The majority of patients undergoing treatment for adenoma or advanced
adenoma were not treated surgically {[}Table 1{]} but rather via
colonoscopy. All patients diagnosed with carcinomas were treated with at
least surgery or a combination of surgery and chemotherapy or surgery,
chemotherapy, and radiation. The type of chemotherapy used for patients
with CRC included Oxaliplatin, Levicovorin, Folfox, Xeloda,
Capecitabine, Avastin, Fluorouracil, and Glucovorin. These were used
individually or in combination with others depending on the patient
{[}Table 1{]}. If an individual was treated with radiation they were
also always treated with chemotherapy. Radiation therapy generally used
18 mV photons for treatment.

\textbf{\emph{16S rRNA Gene Sequencing.}} Sequencing was completed as
described by Kozich, et al. (28). DNA extraction used the 96-well Soil
DNA isolation kit (MO BIO Laboratories) and an epMotion 5075 automated
pipetting system (Eppendorf). The V4 variable region was amplified and
the resulting product was split between four sequencing runs with
normal, adenoma, and carcinoma evenly represented on each run. Each
group was randomly assigned to avoid biases based on sample collection
location. The pre and post-treatment samples were sequenced on the same
run.

\textbf{\emph{Sequence Processing.}} The mothur software package
(v1.37.5) was used to process the 16S rRNA gene sequences and has been
previously described (28). The general workflow using mothur included
merging paired-end reads into contigs, filtering for low quality
contigs, aligning to the SILVA database (29), screening for chimeras
using UCHIME (30), classifying with a naive Bayesian classifier using
the Ribosomal Database Project (RDP)(31), and clustered into Operational
Taxonomic Units (OTUs) using a 97\% similarity cutoff with an average
neighbor clustering algorithm (32). The number of sequences for each
sample was rarefied to 10523 to minimize the impacts of uneven sampling.

\textbf{\emph{Model Building.}} The Random Forest (33) algorithm was
used to create the three models used to classify pre and post-treatment
samples by diagnosis (adenoma, advanced adenoma, or carcinoma). The
total number of individuals in the pre versus post-treatment models was
67 individuals. There were a total of 22 individuals in the pre versus
post-treatment adenoma model, 19 individuals in the pre versus
post-treatment advanced adenoma model, and 26 individuals in the pre
versus post-treatment carcinoma model {[}Table 1{]}.

Similarly, the Random Forest (33) algorithm was also used to create the
three models used to classify normal versus diagnosis. These samples
were obtained using the same methodology as described earlier in this
section. All samples used for this component of model training were from
pre-treatment samples only. The total number of individuals in the
normal versus diagnosis models was 423 individuals {[}Table 2{]}. There
were a total of 239 individuals in the normal versus adenoma model, 262
individuals in the normal versus advanced adenoma model, and 266
individuals in the normal versus carcinoma model {[}Table 2{]}.

All models included only OTU data obtained from 16S rRNA sequencing and
were processed and cleaned using the R package caret (v6.0.76).
Optimization of the mtry hyper-parameter involved making 100 different
80/20 (train/test) splits of the data where the same proportion was
present within both the whole data set and the 80/20 split. For each of
the different splits, 20 repeated 10-fold cross validation was performed
on the 80\% component to optimize the mtry hyper-parameter by maximizing
the AUC (Area Under the Curve of the Receiver Operator Characteristic).
The resulting model was then tested on the hold out data obtained from
the 20\% component. For adenoma pre versus post-treatment models the
optimized mtrys were 2 (n = 42), 42 (n = 26), and 914 (n = 32) for
adenoma. For advanced adenoma the mtrys were 2 (n = 26), 52 (n = 14),
and 1353 (n = 60). For carcinoma the mtrys were 2 (n = 36), 41 (n = 54),
and 863 (n = 10). For the normal versus diagnosis models the optimized
mtrys were 2 (n = 16), 37 (n = 44), and 707 (n = 40) for adenoma. For
advanced adenoma 2 (n = 33), 37 (n = 50), and 696 (n = 17). For
carcinoma 36 (n = 96) and 649 (n = 4). The hyper-parameter, mtry,
defines the number of variables to investigate at each split before a
new division of the data was created with the Random Forest model (33).

The treatment models were then assessed for their strength to classify
whether individuals had obtained treatment. We also assessed whether the
models performed better than random by creating 100 permutations of a
data set in which the labels were randomly assigned. To assess
similarity of the treatment models we looked for overlap between their
top 10 most important OTUs.

For the diagnosis models that were trained on a completely independent
sample set, each model was then applied to our 67-person cohort {[}Table
1{]} based on diagnosis: adenoma (pre-treatment adenoma (adenoma n = 22
and disease free n = 0) versus post-treatment adenoma (adenoma n = 0 and
disease free n = 22)), advanced adenoma pre-treatment advanced adenoma
(advanced adenoma n = 19 and disease free n = 0 ) versus post-treatment
advanced adenoma (advanced adenoma n = 0 and disease free n = 19), and
carcinoma (pre-treatment carcinoma (carcinoma n = 26 and disease free n
= 0) versus post-treatment carcinoma (carcinoma n = 1 and disease free n
= 25)). The application of the pre versus post-treatment models
generated the probabilities that the sample was a pre-treatment sample.
The application of the normal versus diagnosis models generated the
probabilities that the sample was that specific diagnosis (adenoma,
advanced adenoma, or carcinoma). We also looked at similarities between
the models by looking for overlap in the top 10\% of important OTUs to
each respective model.

\textbf{\emph{Statistical Analysis.}} The R software package (v3.4.1)
was used for all statistical analysis. Comparisons between bacterial
community structure utilized PERMANOVA (34) in the vegan package
(v2.4.3). Comparisons between probabilities as well as overall
differences in the median relative abundance of each OTU between pre and
post-treatment samples utilized a paired Wilcoxon ranked sum test.
Comparisons between different treatment for lesions utilized a Kruskal
Wallis test. Where multiple comparison testing was appropriate, a
Benjamini-Hochberg (BH) correction was applied (35) and a corrected
P-value of less than 0.05 was considered significant. The P-values
reported are those that were BH corrected. Model rank importance was
determined by obtaining the median MDA from the 100, 20 repeated 10-fold
cross validation and then ranking from largest to smallest MDA.

\textbf{\emph{Reproducible Methods.}} A detailed and reproducible
description of how the data were processed and analyzed can be found at
\url{https://github.com/SchlossLab/Sze_FollowUps_Microbiome_2017}. Raw
sequences have been deposited into the NCBI Sequence Read Archive
(SRP062005 and SRP096978) and the necessary metadata can be found at
\url{https://www.ncbi.nlm.nih.gov/Traces/study/} and searching the
respective SRA study accession.

\newpage

\subsection{Declarations}\label{declarations}

\subsubsection{Ethics approval and consent to
participate}\label{ethics-approval-and-consent-to-participate}

The University of Michigan Institutional Review Board approved this
study, and all subjects provided informed consent. This study conformed
to the guidelines of the Helsinki Declaration.

\subsubsection{Consent for publication}\label{consent-for-publication}

Not applicable.

\subsubsection{Availability of data and
material}\label{availability-of-data-and-material}

A detailed and reproducible description of how the data were processed
and analyzed can be found at
\url{https://github.com/SchlossLab/Sze_followUps_2017}. Raw sequences
have been deposited into the NCBI Sequence Read Archive (SRP062005 and
SRP096978) and the necessary metadata can be found at
\url{https://www.ncbi.nlm.nih.gov/Traces/study/} and searching the
respective SRA study accession.

\subsubsection{Competing Interests}\label{competing-interests}

All authors declare that they do not have any relevant competing
interests to report.

\subsubsection{Funding}\label{funding}

This study was supported by funding from the National Institutes of
Health to P. Schloss (R01GM099514, P30DK034933) and to the Early
Detection Research Network (U01CA86400).

\subsubsection{Authors' contributions}\label{authors-contributions}

All authors were involved in the conception and design of the study. MAS
analyzed the data. NTB processed samples and analyzed the data. All
authors interpreted the data. MAS and PDS wrote the manuscript. All
authors reviewed and revised the manuscript. All authors read and
approved the final manuscript.

\subsubsection{Acknowledgements}\label{acknowledgements}

The authors thank the Great Lakes-New England Early Detection Research
Network for providing the fecal samples that were used in this study. We
would also like to thank Amanda Elmore for reviewing and correcting code
error and providing feedback on manuscript drafts. We would also like to
thank Nicholas Lesniak for providing feedback on manuscript drafts.

\newpage

\subsection{References}\label{references}

\hypertarget{refs}{}
\hypertarget{ref-siegel_cancer_2016}{}
1. Siegel RL, Miller KD, Jemal A. 2016. Cancer statistics, 2016. CA: a
cancer journal for clinicians 66:7--30.

\hypertarget{ref-haggar_colorectal_2009}{}
2. Haggar FA, Boushey RP. 2009. Colorectal cancer epidemiology:
Incidence, mortality, survival, and risk factors. Clinics in Colon and
Rectal Surgery 22:191--197.

\hypertarget{ref-hellinger_reoperation_2006}{}
3. Hellinger MD, Santiago CA. 2006. Reoperation for recurrent colorectal
cancer. Clinics in Colon and Rectal Surgery 19:228--236.

\hypertarget{ref-ryuk_predictive_2014}{}
4. Ryuk JP, Choi G-S, Park JS, Kim HJ, Park SY, Yoon GS, Jun SH, Kwon
YC. 2014. Predictive factors and the prognosis of recurrence of
colorectal cancer within 2 years after curative resection. Annals of
Surgical Treatment and Research 86:143--151.

\hypertarget{ref-national_cancer_institute_seer_nodate}{}
5. Institute NC. SEER Cancer Stat Facts: Colon and Rectum Cancer.

\hypertarget{ref-goodwin_polyamine_2011}{}
6. Goodwin AC, Destefano Shields CE, Wu S, Huso DL, Wu X, Murray-Stewart
TR, Hacker-Prietz A, Rabizadeh S, Woster PM, Sears CL, Casero RA. 2011.
Polyamine catabolism contributes to enterotoxigenic Bacteroides
fragilis-induced colon tumorigenesis. Proceedings of the National
Academy of Sciences of the United States of America 108:15354--15359.

\hypertarget{ref-abed_fap2_2016}{}
7. Abed J, Emgård JEM, Zamir G, Faroja M, Almogy G, Grenov A, Sol A,
Naor R, Pikarsky E, Atlan KA, Mellul A, Chaushu S, Manson AL, Earl AM,
Ou N, Brennan CA, Garrett WS, Bachrach G. 2016. Fap2 Mediates
Fusobacterium nucleatum Colorectal Adenocarcinoma Enrichment by Binding
to Tumor-Expressed Gal-GalNAc. Cell Host \& Microbe 20:215--225.

\hypertarget{ref-arthur_microbial_2014}{}
8. Arthur JC, Gharaibeh RZ, Mühlbauer M, Perez-Chanona E, Uronis JM,
McCafferty J, Fodor AA, Jobin C. 2014. Microbial genomic analysis
reveals the essential role of inflammation in bacteria-induced
colorectal cancer. Nature Communications 5:4724.

\hypertarget{ref-kostic_fusobacterium_2013}{}
9. Kostic AD, Chun E, Robertson L, Glickman JN, Gallini CA, Michaud M,
Clancy TE, Chung DC, Lochhead P, Hold GL, El-Omar EM, Brenner D, Fuchs
CS, Meyerson M, Garrett WS. 2013. Fusobacterium nucleatum potentiates
intestinal tumorigenesis and modulates the tumor-immune
microenvironment. Cell Host \& Microbe 14:207--215.

\hypertarget{ref-wu_human_2009}{}
10. Wu S, Rhee K-J, Albesiano E, Rabizadeh S, Wu X, Yen H-R, Huso DL,
Brancati FL, Wick E, McAllister F, Housseau F, Pardoll DM, Sears CL.
2009. A human colonic commensal promotes colon tumorigenesis via
activation of T helper type 17 T cell responses. Nature Medicine
15:1016--1022.

\hypertarget{ref-zackular_manipulation_2016}{}
11. Zackular JP, Baxter NT, Chen GY, Schloss PD. 2016. Manipulation of
the Gut Microbiota Reveals Role in Colon Tumorigenesis. mSphere 1.

\hypertarget{ref-zackular_gut_2013}{}
12. Zackular JP, Baxter NT, Iverson KD, Sadler WD, Petrosino JF, Chen
GY, Schloss PD. 2013. The gut microbiome modulates colon tumorigenesis.
mBio 4:e00692--00613.

\hypertarget{ref-baxter_structure_2014}{}
13. Baxter NT, Zackular JP, Chen GY, Schloss PD. 2014. Structure of the
gut microbiome following colonization with human feces determines
colonic tumor burden. Microbiome 2:20.

\hypertarget{ref-flynn_metabolic_2016}{}
14. Flynn KJ, Baxter NT, Schloss PD. 2016. Metabolic and Community
Synergy of Oral Bacteria in Colorectal Cancer. mSphere 1.

\hypertarget{ref-wang_structural_2012}{}
15. Wang T, Cai G, Qiu Y, Fei N, Zhang M, Pang X, Jia W, Cai S, Zhao L.
2012. Structural segregation of gut microbiota between colorectal cancer
patients and healthy volunteers. The ISME journal 6:320--329.

\hypertarget{ref-chen_decreased_2013}{}
16. Chen H-M, Yu Y-N, Wang J-L, Lin Y-W, Kong X, Yang C-Q, Yang L, Liu
Z-J, Yuan Y-Z, Liu F, Wu J-X, Zhong L, Fang D-C, Zou W, Fang J-Y. 2013.
Decreased dietary fiber intake and structural alteration of gut
microbiota in patients with advanced colorectal adenoma. The American
Journal of Clinical Nutrition 97:1044--1052.

\hypertarget{ref-chen_human_2012}{}
17. Chen W, Liu F, Ling Z, Tong X, Xiang C. 2012. Human intestinal lumen
and mucosa-associated microbiota in patients with colorectal cancer.
PloS One 7:e39743.

\hypertarget{ref-shen_molecular_2010}{}
18. Shen XJ, Rawls JF, Randall T, Burcal L, Mpande CN, Jenkins N, Jovov
B, Abdo Z, Sandler RS, Keku TO. 2010. Molecular characterization of
mucosal adherent bacteria and associations with colorectal adenomas. Gut
Microbes 1:138--147.

\hypertarget{ref-kostic_genomic_2012}{}
19. Kostic AD, Gevers D, Pedamallu CS, Michaud M, Duke F, Earl AM,
Ojesina AI, Jung J, Bass AJ, Tabernero J, Baselga J, Liu C, Shivdasani
RA, Ogino S, Birren BW, Huttenhower C, Garrett WS, Meyerson M. 2012.
Genomic analysis identifies association of Fusobacterium with colorectal
carcinoma. Genome Research 22:292--298.

\hypertarget{ref-feng_gut_2015}{}
20. Feng Q, Liang S, Jia H, Stadlmayr A, Tang L, Lan Z, Zhang D, Xia H,
Xu X, Jie Z, Su L, Li X, Li X, Li J, Xiao L, Huber-Schönauer U,
Niederseer D, Xu X, Al-Aama JY, Yang H, Wang J, Kristiansen K, Arumugam
M, Tilg H, Datz C, Wang J. 2015. Gut microbiome development along the
colorectal adenoma-carcinoma sequence. Nature Communications 6:6528.

\hypertarget{ref-dejea_microbiota_2014}{}
21. Dejea CM, Wick EC, Hechenbleikner EM, White JR, Mark Welch JL,
Rossetti BJ, Peterson SN, Snesrud EC, Borisy GG, Lazarev M, Stein E,
Vadivelu J, Roslani AC, Malik AA, Wanyiri JW, Goh KL, Thevambiga I, Fu
K, Wan F, Llosa N, Housseau F, Romans K, Wu X, McAllister FM, Wu S,
Vogelstein B, Kinzler KW, Pardoll DM, Sears CL. 2014. Microbiota
organization is a distinct feature of proximal colorectal cancers.
Proceedings of the National Academy of Sciences of the United States of
America 111:18321--18326.

\hypertarget{ref-mima_fusobacterium_2015}{}
22. Mima K, Sukawa Y, Nishihara R, Qian ZR, Yamauchi M, Inamura K, Kim
SA, Masuda A, Nowak JA, Nosho K, Kostic AD, Giannakis M, Watanabe H,
Bullman S, Milner DA, Harris CC, Giovannucci E, Garraway LA, Freeman GJ,
Dranoff G, Chan AT, Garrett WS, Huttenhower C, Fuchs CS, Ogino S. 2015.
Fusobacterium nucleatum and T Cells in Colorectal Carcinoma. JAMA
oncology 1:653--661.

\hypertarget{ref-arthur_intestinal_2012}{}
23. Arthur JC, Perez-Chanona E, Mühlbauer M, Tomkovich S, Uronis JM, Fan
T-J, Campbell BJ, Abujamel T, Dogan B, Rogers AB, Rhodes JM, Stintzi A,
Simpson KW, Hansen JJ, Keku TO, Fodor AA, Jobin C. 2012. Intestinal
inflammation targets cancer-inducing activity of the microbiota. Science
(New York, NY) 338:120--123.

\hypertarget{ref-baxter_microbiota-based_2016}{}
24. Baxter NT, Ruffin MT, Rogers MAM, Schloss PD. 2016. Microbiota-based
model improves the sensitivity of fecal immunochemical test for
detecting colonic lesions. Genome Medicine 8:37.

\hypertarget{ref-zeller_potential_2014}{}
25. Zeller G, Tap J, Voigt AY, Sunagawa S, Kultima JR, Costea PI, Amiot
A, Böhm J, Brunetti F, Habermann N, Hercog R, Koch M, Luciani A, Mende
DR, Schneider MA, Schrotz-King P, Tournigand C, Tran Van Nhieu J, Yamada
T, Zimmermann J, Benes V, Kloor M, Ulrich CM, Knebel Doeberitz M von,
Sobhani I, Bork P. 2014. Potential of fecal microbiota for early-stage
detection of colorectal cancer. Molecular Systems Biology 10:766.

\hypertarget{ref-zackular_human_2014}{}
26. Zackular JP, Rogers MAM, Ruffin MT, Schloss PD. 2014. The human gut
microbiome as a screening tool for colorectal cancer. Cancer Prevention
Research (Philadelphia, Pa) 7:1112--1121.

\hypertarget{ref-obrien_impact_2013}{}
27. O'Brien CL, Allison GE, Grimpen F, Pavli P. 2013. Impact of
colonoscopy bowel preparation on intestinal microbiota. PloS One
8:e62815.

\hypertarget{ref-kozich_development_2013}{}
28. Kozich JJ, Westcott SL, Baxter NT, Highlander SK, Schloss PD. 2013.
Development of a dual-index sequencing strategy and curation pipeline
for analyzing amplicon sequence data on the MiSeq Illumina sequencing
platform. Applied and Environmental Microbiology 79:5112--5120.

\hypertarget{ref-pruesse_silva_2007}{}
29. Pruesse E, Quast C, Knittel K, Fuchs BM, Ludwig W, Peplies J,
Glöckner FO. 2007. SILVA: A comprehensive online resource for quality
checked and aligned ribosomal RNA sequence data compatible with ARB.
Nucleic Acids Research 35:7188--7196.

\hypertarget{ref-edgar_uchime_2011}{}
30. Edgar RC, Haas BJ, Clemente JC, Quince C, Knight R. 2011. UCHIME
improves sensitivity and speed of chimera detection. Bioinformatics
(Oxford, England) 27:2194--2200.

\hypertarget{ref-wang_naive_2007}{}
31. Wang Q, Garrity GM, Tiedje JM, Cole JR. 2007. Naive Bayesian
classifier for rapid assignment of rRNA sequences into the new bacterial
taxonomy. Applied and Environmental Microbiology 73:5261--5267.

\hypertarget{ref-schloss_assessing_2011}{}
32. Schloss PD, Westcott SL. 2011. Assessing and improving methods used
in operational taxonomic unit-based approaches for 16S rRNA gene
sequence analysis. Applied and Environmental Microbiology 77:3219--3226.

\hypertarget{ref-breiman_random_2001}{}
33. Breiman L. 2001. Random Forests. Machine Learning 45:5--32.

\hypertarget{ref-anderson_permanova_2013}{}
34. Anderson MJ, Walsh DCI. 2013. PERMANOVA, ANOSIM, and the Mantel test
in the face of heterogeneous dispersions: What null hypothesis are you
testing? Ecological Monographs 83:557--574.

\hypertarget{ref-benjamini_controlling_1995}{}
35. Benjamini Y, Hochberg Y. 1995. Controlling the false discovery rate:
A practical and powerful approach to multiple testing. Journal of the
Royal Statistical Society Series B (Methodological) 57:289--300.

\newpage

\textbf{Table 1: Demographic data of patients in the pre and
post-treatment cohort}

\begin{longtable}[]{@{}lccc@{}}
\toprule
& Adenoma & Advanced Adenoma & Carcinoma\tabularnewline
\midrule
\endhead
n & 22 & 19 & 26\tabularnewline
Age (Mean ± SD) & 61.68 ± 7.2 & 63.11 ± 10.9 & 61.65 ±
12.9\tabularnewline
Sex (\%F) & 36.36 & 36.84 & 42.31\tabularnewline
BMI (Mean ± SD) & 26.86 ± 3.9 & 25.81 ± 4.7 & 28.63 ± 7.2\tabularnewline
Caucasian (\%) & 95.45 & 84.21 & 96.15\tabularnewline
Days Between Colonoscopy (Mean ± SD) & 255.41 ± 42 & 250.16 ± 41 &
350.85 ± 102\tabularnewline
Surgery Only & 4 & 4 & 12\tabularnewline
Surgery \& Chemotherapy & 0 & 0 & 9\tabularnewline
Surgery, Chemotherapy, \& Radiation & 0 & 0 & 5\tabularnewline
\bottomrule
\end{longtable}

\newpage

\textbf{Table 2: Demographic data of training cohort}

\begin{longtable}[]{@{}lcccc@{}}
\toprule
& Normal & Adenoma & Advanced Adenoma & Carcinoma\tabularnewline
\midrule
\endhead
n & 172 & 67 & 90 & 94\tabularnewline
Age (Mean ± SD) & 54.29 ± 9.9 & 63.01 ± 13.1 & 64.07 ± 11.3 & 64.37 ±
12.9\tabularnewline
Sex (\%F) & 64.53 & 46.27 & 37.78 & 43.62\tabularnewline
BMI (Mean ± SD) & 26.97 ± 5.3 & 25.69 ± 4.8 & 26.66 ± 4.9 & 29.27 ±
6.7\tabularnewline
Caucasian (\%) & 87.79 & 92.54 & 92.22 & 94.68\tabularnewline
\bottomrule
\end{longtable}

\newpage

\textbf{Figure 1: General differences between adenoma, advanced adenoma,
and carcinoma groups after treatment.} A) \(\theta\)\textsubscript{YC}
distances from pre versus post sample within each individual. A
significant difference was found between the adenoma and carcinoma group
(P-value = 5.36e-05). Solid black points represent the median value for
each diagnosis group. B) NMDS of the pre and post-treatment samples for
the adenoma group. C) NMDS of the pre and post-treatment samples for the
advanced adenoma group. D) NMDS of the pre and post-treatment samples
for the carcinoma group.

\textbf{Figure 2: The top 10 most important OTUs used to classify
treatment for adenoma, advanced adenoma, and carcinoma.} A) Adenoma
OTUs. B) Advanced Adenoma OTUs. C) Carcinoma OTUs. The darker circle
highlights the median log10 MDA value obtained from 100 different 80/20
splits while the lighter colored circles represents the value obtained
for a specific run.

\textbf{Figure 3: Top 10\% most important OTUs common to those models
used to differentiate between patients with normal colons and those with
adenoma, advanced adenoma, and carcinoma.} A) Venn diagram showing the
OTU overlap between each model. B) For each common OTU the lowest
taxonomic identification and importance rank for each model run is
shown.

\textbf{Figure 4: Treatment response based on models built for adenoma,
advanced adenoma, or carcinoma.} A) Positive probability change from
initial to follow up sample in those with adenoma. B) Positive
probability change from initial to follow up sample in those with
advanced adenoma. C) Positive probability change from initial to follow
up sample in those with carcinoma.

\newpage

\textbf{Figure S1: ROC curves of the adenoma, advanced adenoma, and
carcinoma models.} A) Adenoma ROC curve: The light green shaded areas
represent the range of values of a 100 different 80/20 splits of the
test set data and the dark green line represents the model using 100\%
of the data set and what was used for subsequent classification. B)
Advanced Adenoma ROC curve: The light yellow shaded areas represent the
range of values of a 100 different 80/20 splits of the test set data and
the dark yellow line represents the model using 100\% of the data set
and what was used for subsequent classification. C) Carcinoma ROC curve:
The light red shaded areas represent the range of values of a 100
different 80/20 splits of the test set data and the dark red line
represents the model using 100\% of the data set and what was used for
subsequent classification.

\textbf{Figure S2: Summary of top 10\% of important OTUs for the
adenoma, advanced adenoma, and carcinoma models.} A) MDA of the most
important variables in the adenoma model. The dark green point
represents the mean and the lighter green points are the value of each
of the 100 different runs. B) Summary of Important Variables in the
advanced adenoma model. MDA of the most important variables in the SRN
model. The dark yellow point represents the mean and the lighter yellow
points are the value of each of the 100 different runs. C) MDA of the
most important variables in the carcinoma model. The dark red point
represents the mean and the lighter red points are the value of each of
the 100 different runs.

\textbf{Figure S3: Pre and post-treatment relative abundance of CRC
associated OTUs within the carcinoma model.}

\newpage

\end{document}